---
title: "north-parcel-experiment"
author: "Joanna Tang"
date: "11/14/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}

# Loading packages, read in data
library(tidyverse)
library(janitor)
library(knitr)
library(kableExtra)
np_2016_seedbank_master <- read_csv("np_2016_seedbank.csv")
metadata <- read_csv("metadata.csv")

```


```{r echo = FALSE, message = FALSE, warning = FALSE}

# Convert np_2016_seedbank into tidy format, remove NAs, separate "Replicate" label into Replicate_Zone and Replicate_Number
seedbank_2016 <- gather(np_2016_seedbank_master, key = "Species", value = "Count", -c(Plot, Replicate)) %>% 
  na.omit(np_2016_seedbank_master) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3"))


# Make separate dataframes for each site, with % column and species metadata
ph1 <- filter(seedbank_2016, Plot == "PH1") %>% 
  mutate(Percent = 100*Count/sum(Count))
ph1_2016_seedbank <- full_join(ph1, metadata)

ms <- filter(seedbank_2016, Plot == "MS") %>% 
  mutate(Percent = 100*Count/sum(Count))
ms_2016_seedbank <- full_join(ph1_2016_seedbank, metadata)

rt <- filter(seedbank_2016, Plot == "RT") %>% 
  mutate(Percent = 100*Count/sum(Count))
rt_2016_seedbank <- full_join(rt, metadata)

tp <- filter(seedbank_2016, Plot == "TP") %>% 
  mutate(Percent = 100*Count/sum(Count))
tp_2016_seedbank <- full_join(tp, metadata)

wt1 <- filter(seedbank_2016, Plot == "WT1") %>% 
  mutate(Percent = 100*Count/sum(Count))
wt1_2016_seedbank <- full_join(wt1, metadata)

wt2 <- filter(seedbank_2016, Plot == "WT2") %>% 
  mutate(Percent = 100*Count/sum(Count))
wt2_2016_seedbank <- full_join(wt2, metadata)

cs <- filter(seedbank_2016, Plot == "CS") %>% 
  mutate(Percent = 100*Count/sum(Count))
cs_2016_seedbank <- full_join(cs, metadata)


```

# Phase 1

PSBR present, JUBU dominant

Exotics increase from C to T to U

```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate PH1 data
ph1_2016_seedbank_summary <- ph1_2016_seedbank %>% 
  na.omit(ph1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
ph1_col <- ggplot(ph1_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
ph1_col

# Central zone percentages
ph1_c_col <- filter(ph1_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(ph1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
ph1_c_col

# Transition zone percentages
ph1_t_col <- filter(ph1_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(ph1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
ph1_t_col

# Upland zone percentages
ph1_u_col <- filter(ph1_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(ph1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
ph1_u_col


# Native vs. exotic
ph1_native <- ggplot(filter(ph1_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
ph1_native

```

# Mini South

PSBR present, JUBU dominant

Exotics increase from C to T to U

```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate MS data
ms_2016_seedbank_summary <- ms_2016_seedbank %>% 
  na.omit(ms_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
ms_col <- ggplot(ms_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
ms_col

# Central zone percentages
ms_c_col <- filter(ms_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(ms_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
ms_c_col

# Transition zone percentages
ms_t_col <- filter(ms_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(ms_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
ms_t_col

# Upland zone percentages
ms_u_col <- filter(ms_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(ms_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
ms_u_col


# Native vs. exotic
ms_native <- ggplot(filter(ms_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
ms_native

```

# Redtail

JUBU dominant

Exotics increase from C to T

```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate RT data
rt_2016_seedbank_summary <- rt_2016_seedbank %>% 
  na.omit(rt_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
rt_col <- ggplot(rt_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
rt_col

# Central zone percentages
rt_c_col <- filter(rt_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(rt_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
rt_c_col

# Transition zone percentages
rt_t_col <- filter(rt_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(rt_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
rt_t_col

# Upland zone percentages
rt_u_col <- filter(rt_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(rt_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
rt_u_col


# Native vs. exotic
rt_native <- ggplot(filter(rt_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
rt_native

```

# Tadpole

ELMA present, JUBU dominant

Exotics increase from C to T to U

```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate TP data
tp_2016_seedbank_summary <- tp_2016_seedbank %>% 
  na.omit(tp_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
tp_col <- ggplot(tp_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
tp_col

# Central zone percentages
tp_c_col <- filter(tp_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(rt_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
tp_c_col

# Transition zone percentages
tp_t_col <- filter(tp_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(tp_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
tp_t_col

# Upland zone percentages
tp_u_col <- filter(tp_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(tp_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
tp_u_col



# Native vs. exotic
tp_native <- ggplot(filter(tp_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
tp_native

```

# Whitetail 1

JUBU dominant

Exotics increase from C to T, dominate in C
```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate WT1 data
wt1_2016_seedbank_summary <- wt1_2016_seedbank %>% 
  na.omit(rt_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
wt1_col <- ggplot(wt1_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
wt1_col

# Central zone percentages
wt1_c_col <- filter(wt1_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(wt1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
wt1_c_col

# Transition zone percentages
wt1_t_col <- filter(wt1_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(wt1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
wt1_t_col

# Upland zone percentages
wt1_u_col <- filter(wt1_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(wt1_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
wt1_u_col



# Native vs. exotic
wt1_native <- ggplot(filter(wt1_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
wt1_native

```

# Whitetail 2

JUBU dominant

Exotics decrease from C to T to U, dominate in C
```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate WT2 data
wt2_2016_seedbank_summary <- wt2_2016_seedbank %>% 
  na.omit(wt2_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
wt2_col <- ggplot(wt2_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
wt2_col

# Central zone percentages
wt2_c_col <- filter(wt2_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(wt2_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
wt2_c_col

# Transition zone percentages
wt2_t_col <- filter(wt2_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(wt2_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
wt2_t_col

# Upland zone percentages
wt2_u_col <- filter(wt2_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(wt2_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
wt2_u_col



# Native vs. exotic
wt2_native <- ggplot(filter(wt2_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
wt2_native

```

# Creekside

PSBR present, JUBU dominant

Exotics fairly constant from C to T to U
```{r echo = FALSE, message = FALSE, warning = FALSE}

# Aggregate CS data
cs_2016_seedbank_summary <- cs_2016_seedbank %>% 
  na.omit(cs_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    )

# Aggregate percentages
cs_col <- ggplot(cs_2016_seedbank_summary, aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate percentages", y = "Mean Percentage")
cs_col

# Central zone percentages
cs_c_col <- filter(cs_2016_seedbank, Replicate_Zone == "C") %>% 
  na.omit(cs_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Central Zone", y = "Mean Percentage")
cs_c_col

# Transition zone percentages
cs_t_col <- filter(cs_2016_seedbank, Replicate_Zone == "T") %>% 
  na.omit(cs_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Transition Zone", y = "Mean Percentage")
cs_t_col

# Upland zone percentages
cs_u_col <- filter(cs_2016_seedbank, Replicate_Zone == "U") %>% 
  na.omit(cs_2016_seedbank) %>% 
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count),
    mean_percentage = mean(Percent)
    ) %>% 
  ggplot(aes(x = Species, y = mean_percentage)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Upland Zone", y = "Mean Percentage")
cs_u_col



# Native vs. exotic
cs_native <- ggplot(filter(cs_2016_seedbank, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
cs_native

```

# Aggregate
```{r echo = FALSE, message = FALSE, warning = FALSE}

seedbank_2016_summary <- full_join(seedbank_2016, metadata) %>% 
  na.omit(seedbank_2016) %>%
  group_by(Species, Native_Status) %>% 
  summarize(
    total = sum(Count)
    )

# Aggregate counts
np_col <- ggplot(seedbank_2016_summary, aes(x = Species, y = total)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate counts", x = "Total Counts", y = "Mean Percentage")
np_col

# Aggregate counts <20
np_col_under_20 <- ggplot(filter(seedbank_2016_summary, total <20), aes(x = Species, y = total)) +
  geom_col(aes(fill = Native_Status)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Aggregate counts <20", x = "Total Counts", y = "Mean Percentage")
np_col_under_20




# Native vs. exotic
np_native <- full_join(seedbank_2016, metadata)
np_native <- ggplot(filter(np_native, Native_Status == "N" | Native_Status == "E"), aes(x = Native_Status)) +
  geom_bar() +
  labs(title = "Native vs. exotic", x = "Native Status", y = "Frequency")
np_native

```

