---
title: "NP Draft 7"
author: "Joanna Tang"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

# Loading packages, read in data  
library(pairwiseAdonis)
library(kableExtra)
library(vegan)
library(car)
library(RColorBrewer)
library(reshape2)
library(kableExtra)
library(FSA)
library(plotly)
library(ggsignif)
library(lme4)
library(lmerTest)
library(emmeans)
library(bestNormalize)
library(betareg)
library(nlme)
library(MASS)
library(fitdistrplus)
library(plotrix)
library(janitor)
library(ggpubr)
library(rstatix)
library(NbClust)
library(betareg)
library(glmmTMB)
library(tidyverse)


np_percent_cover_master <- read_csv("np_percent_cover.csv")
metadata <- read_csv("metadata.csv")
np_2016_seedbank_master <- read_csv("np_2016_seedbank.csv")
np_2018_seed_bank_master <- read_csv("np.2018.seed.bank.csv")
del_sol_seedbank_master <- read_csv("del_sol_seedbank.csv")
biomass_master <- read_csv("combined_biomass.csv")
del_sol_master <- read_csv("delsol.csv")
noaa <- read_csv("1567450.csv")
np_pc_ord <- read_csv("pc-ord_nms_scores.csv")
np_community_matrix_variables <- read_csv("np_community_matrix_variables.csv")
hydro <- read_csv("2019_np_hydro_size.csv")


```


```{r include = FALSE}

# Outline

# 0. Tidy up data

# 1. Is sustained weeding effort a successful long-term restoration strategy?
## 1a. How does total exotic percent cover, biomass, and diversity change with age of pool? (boxplots)
## 1b. Do pools converge on the same exotic community composition over time? (NMDS, cluster)
##1c. Is the whole community changing? (NMDS, cluster)
##1d. Is the native community changing? (NMDS, cluster, column graphs)


```

```{r include = FALSE}

# 0. Tidy up data

## Create new data frame with creation year of each pool
creation_year_df <- data.frame(Plot = c("PH1", "TP", "RT", "MS", "WT1", "WT2", "CS"), Creation_Year = c("2010", "2012", "2012", "2015", "2014", "2014", "2013"))

## Separate "Replicate" label into Replicate_Zone and Replicate_Number, join with species metadata, convert month and year to "date" column
np_percent_cover <- np_percent_cover_master %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3"))
np_percent_cover <- full_join(np_percent_cover, metadata) %>% 
  mutate(month_number = case_when(Month == "JANUARY" ~ 1, Month == "FEBRUARY" ~ 2, Month == "MARCH" ~ 3, Month == "APRIL" ~ 4, Month == "MAY" ~ 5, Month == "JUNE" ~ 6, Month == "JULY" ~ 7, Month == "AUGUST" ~ 8, Month == "SEPTEMBER" ~ 9, Month == "OCTOBER" ~ 10, Month == "NOVEMBER" ~ 11, Month == "DECEMBER" ~ 12)) %>% 
  mutate(date = paste(Year, month_number, "1", sep="-" )) %>% 
  dplyr::select(-X8)
np_percent_cover$date = as.Date(np_percent_cover$date)


## Convert np_2016_seedbank into tidy format, remove NAs, separate "Replicate" label into Replicate_Zone and Replicate_Number, add Year column, add seedbank column, join to metadata
seedbank_2016 <- gather(np_2016_seedbank_master, key = "Species", value = "Count", -c(Plot, Replicate)) %>% 
  na.omit(np_2016_seedbank_master) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3")) %>% 
  filter(Count>0) %>%
  add_column(Year = 2016) %>% 
  add_column(seedbank = "seedbank_2016") %>% 
  full_join(metadata)

## Calculate total abundance of each species per Replicate for North Parcel 2018 seed bank data, separate "Replicate" label into Replicate_Zone and Replicate_Number, add Year column, add seedbank column, join to metadata
np_seedbank_2018 <- np_2018_seed_bank_master %>% 
  group_by(Plot, Replicate, Species) %>% 
  summarize(Count = sum(Count)) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3")) %>% 
  add_column(Year = 2018) %>% 
  add_column(seedbank = "seedbank_2018") %>% 
  full_join(metadata)

## Calculate total abundance of each species per Replicate for Del Sol 2018 seed bank data, separate "Replicate" label into Replicate_Zone and Replicate_Number, add Year column, add seedbank column, join to metadata
del_sol_seedbank <- del_sol_seedbank_master %>% 
  group_by(Plot, Replicate, Species) %>% 
  summarize(Count = sum(Count)) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3")) %>% 
  add_column(Year = 2018) %>% 
  add_column(seedbank = "del_sol_seedbank_2018") %>% 
  full_join(metadata)

## Join Del Sol % cover data to metadata
del_sol <- full_join(del_sol_master, metadata)


## Join biomass to metadata
annual_exotic_biomass <- biomass_master %>% 
  filter(Year>2016) %>% 
  filter(Year<2020) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  drop_na(Time_Since) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  drop_na(Replicate_Zone) %>% 
  group_by(Year, Plot, Replicate_Zone, Replicate, Creation_Year, Time_Since) %>% 
  summarize(annual_biomass = sum(biomass_g)) %>% 
  mutate(plot_replicate = paste(Plot, Replicate, sep = "-"))



```


# 1. Is sustained weeding effort a successful long-term restoration strategy?

## 1a. How does total exotic percent cover, biomass, and diversity change with age of pool?

Hypothesis: If sustained weeding (routine weed-whacking, monthly hand-weeding) is successful, total exotic cover will remain low.

Results: Significant increase in exotic H' in upland zone at year 4.

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1a. How does total exotic percent cover, biomass, and diversity change with age of pool? (boxplots)

## Create new data frame with creation year of each pool
creation_year_df <- data.frame(Plot = c("PH1", "TP", "RT", "MS", "WT1", "WT2", "CS"), Creation_Year = c("2010", "2012", "2012", "2015", "2014", "2014", "2013"))

## Calculate total exotic % cover per quadrat, find max total exotic % cover per quadrat per year, add time since restoration column, add depth (from max 2019 hydro)
np_annual_total_exotic <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  group_by(Year, Plot, Replicate, Replicate_Zone) %>% 
  summarize(max_annual_exotic_pc = max(total_exotic_pc)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  mutate(plot_replicate = paste(Plot, Replicate, sep = "-")) %>% 
  full_join(hydro)

### Check for normality
exotic_pc_hist <- ggplot(np_annual_total_exotic, aes(x = max_annual_exotic_pc)) +
  geom_histogram()
#not normal -- skewed right
exotic_pc_qq <- ggplot(np_annual_total_exotic, aes(sample = max_annual_exotic_pc)) +
  geom_qq()
#not normal -- skewed right

## Boxplot of total exotic % cover, by zone, by time since
exotic_pc_time_box <- np_annual_total_exotic %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "Central",  Replicate_Zone == "T" ~ "Transition", Replicate_Zone == "U" ~ "Upland")) %>%
  ggplot(aes(x = as.factor(Time_Since), y = max_annual_exotic_pc)) +
  geom_boxplot(fill = "coral3") +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Total Percent Cover of Exotics, over time", x = "Age of Pool (years)", y = "Percent Cover (%)", caption = "Sig increase around 5 in transition and upland \n (Sig increase in transition and upland over time)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100))
exotic_pc_time_box

### LMER for exotic % cover (log-transformed)
exotic_pc_lmer <- lmer(log(max_annual_exotic_pc) ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|plot_replicate), np_annual_total_exotic)
#summary(exotic_pc_lmer)
#some age, age:upland and age:transition sig
#### Check diagnostic graphs
#E1 <- resid(exotic_pc_lmer)
#F1 <- fitted(exotic_pc_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_pc_lmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_pc_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_pc_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_pc_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal
#### ANOVA of lmer
exotic_pc_lmer_anova <- anova(exotic_pc_lmer)
#### Tukey's post-hoc of least-squares means
exotic_pc_lmer_tukey <- emmeans(exotic_pc_lmer, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(exotic_pc_lmer_tukey, comparisons = TRUE)
#Central: no sig
#Transition: 3-5 p = 0.0012, 3-6 p - 0.0242, 3-7 p = -.0013, 3-8 p = 00002, 3-9 p = 0.0113, 4-5 p = 0.0958, 4-7 p = 0.0293, 4-8 p = 0.0023, 4-9 p = 0.0830
#Upland: 3-7 p = 0.0933, 3-5 p = 0.0111, 3-6 p = 0.0285, 3-7 p = 0.0015, 3-8 p = 0.0225, 3-9 p = 0.0204, 4-5 p = 0.0406, 4-6 p = 0.0482, 4-7 p = 0.0024, 4-8 p = 0.0467, 4-9 p = 0.0357
exotic_pc_continuous_lmer <- np_annual_total_exotic %>% 
  filter(Replicate_Zone == "U") %>% 
  lmer(log(max_annual_exotic_pc) ~ (Time_Since) + (1|Year) + (1|plot_replicate), data = .)
#summary(exotic_pc_continuous_lmer)
#transition, upland sig
#### Check diagnostic graphs
#E1 <- resid(exotic_pc_continuous_lmer)
#F1 <- fitted(exotic_pc_continuous_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_pc_continuous_lmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_pc_continuous_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_pc_continuous_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_pc_continuous_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal


#### GLMER for exotic % cover
exotic_pc_glmer <- glmer.nb((max_annual_exotic_pc) ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|depth) + (1|size) + (1|plot_replicate), np_annual_total_exotic)
#summary(exotic_pc_glmer)
#time, transition, upland sig
#### Check diagnostic graphs
#E1 <- resid(exotic_pc_glmer)
#F1 <- fitted(exotic_pc_glmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_pc_glmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_pc_glmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_pc_glmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_pc_glmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_pc_glmer)$depth[,1], main = "depth Residuals")
#qqline(ranef(exotic_pc_glmer)$depth[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_pc_glmer)$size[,1], main = "size Residuals")
#qqline(ranef(exotic_pc_glmer)$size[,1]) #run 2 lines together -- assess if random effects residuals are normal
#Checking for overdispersion -- how well your chosen link distribution matches your data
#E1<-resid(exotic_pc_glmer, type = "pearson")
#N <- length(E1)
#p <- length(coef(exotic_pc_glmer))
#D <- sum(E1^2) / (N - p)
#D = 0.674 -- want this to be 1.0-1.1; above 1.2 is not great; above 1.3 is really not good (overdispersed = easier to accept false sig differences); if underdispersed (<1) then there's less error than the distribution can handle (harder to detect sig diff)
#### ANOVA of glmer
exotic_pc_glmer_anova <- anova(exotic_pc_glmer)
#### Tukey's post-hoc of least-squares means
exotic_pc_glmer_tukey <- emmeans(exotic_pc_glmer, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(exotic_pc_lmer_tukey)






### Check annual total biomass for normality
biomass_hist <- ggplot(annual_exotic_biomass, aes(x = annual_biomass)) +
  geom_histogram()
#not normal -- skewed right
biomass_qq <- ggplot(annual_exotic_biomass, aes(sample = annual_biomass)) +
  geom_qq()
#not normal -- skewed right

## Boxplot of total annual exotic biomass, by zone, by time since
exotic_biomass_time_box <- annual_exotic_biomass %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "Central",  Replicate_Zone == "T" ~ "Transition", Replicate_Zone == "U" ~ "Upland")) %>%
  ggplot(aes(x = as.factor(Time_Since), y = annual_biomass)) +
  geom_boxplot(fill = "coral3") +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Total Biomass of Exotics, over time", x = "Age of Pool (years)", y = "Biomass (g)", caption = "Sig increase around 4 in transition and upland \n (Sig increase in transition and upland over time)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100))
exotic_biomass_time_box

### LMER (log+1 transformed)
exotic_biomass_lmer <- lmer(log(annual_biomass+1) ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|plot_replicate), annual_exotic_biomass)
#summary(exotic_biomass_lmer)
#some age, age:upland and age:transition sig
#### Check diagnostic graphs
#E1 <- resid(exotic_biomass_lmer)
#F1 <- fitted(exotic_biomass_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_biomass_lmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_biomass_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_biomass_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_biomass_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal
#### ANOVA of lmer
exotic_biomass_lmer_anova <- anova(exotic_biomass_lmer)
#### Tukey's post-hoc of least-squares means
exotic_biomass_lmer_tukey <- emmeans(exotic_biomass_lmer, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(exotic_biomass_lmer_tukey, comparisons = TRUE)
#Central: no sig
#Transition: 2-7 p = 0.0675, 2-8 p = 0.0891, 2-9 p = 0.0784, 3-6 p = 0.767, 3-7 p = 0.0012, 3-8 p = 0.0102, 3-9 p = 0.0107, 4-7 p = 0.0474
#Upland: 3-5 p = 0.0763, 3-6 p = 0.0503, 3-7 p = 0.0101, 3-8 p = 0.1715, 3-8 p = 0.0984
exotic_biomass_continuous_lmer <- annual_exotic_biomass %>% 
  filter(Replicate_Zone == "U") %>% 
  lmer(log(annual_biomass+1) ~ (Time_Since) + (1|Year) + (1|plot_replicate), data = .)
#summary(exotic_biomass_continuous_lmer)
#upland and transition sig
#### Check diagnostic graphs
#E1 <- resid(exotic_biomass_continuous_lmer)
#F1 <- fitted(exotic_biomass_continuous_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_biomass_continuous_lmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_biomass_continuous_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_biomass_continuous_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_biomass_continuous_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal


#### GLMER for exotic biomass
exotic_biomass_glmer <- full_join(annual_exotic_biomass, hydro) %>% glmer((annual_biomass+1) ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|depth) + (1|size) + (1|plot_replicate), family = Gamma(link = "log"), data = .)
#summary(exotic_biomass_glmer)
#time, transition, upland sig
#### Check diagnostic graphs
#E1 <- resid(exotic_biomass_glmer)
#F1 <- fitted(exotic_biomass_glmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_biomass_glmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_biomass_glmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_biomass_glmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_biomass_glmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_biomass_glmer)$depth[,1], main = "depth Residuals")
#qqline(ranef(exotic_biomass_glmer)$depth[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_biomass_glmer)$size[,1], main = "Msize Residuals")
#qqline(ranef(exotic_biomass_glmer)$size[,1]) #run 2 lines together -- assess if random effects residuals are normal
#### ANOVA of glmer
exotic_biomass_glmer_anova <- anova(exotic_biomass_glmer)
#### Tukey's post-hoc of least-squares means
exotic_biomass_glmer_tukey <- emmeans(exotic_biomass_glmer, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(exotic_pc_lmer_tukey)









## Shannon's Index for exotic community of each plot per year
exotic_h <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "E") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  group_by(date, month_number, Year, Plot, Replicate, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(8:71) %>% 
  diversity() %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>%
  rename("h" = ".")

## Max H' for each quadrat per zone per year, add time since column
max_annual_exotic_h <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "E") %>% 
  group_by(date, month_number, Year, Plot, Replicate, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  select(1:8) %>% 
  full_join(exotic_h) %>% 
  group_by(Year, Plot, Replicate, Replicate_Zone) %>% 
  mutate(max_exotic_h = max(h)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  mutate(plot_replicate = paste(Plot, Replicate, sep = "-"))

### Exploratory graphs
exotic_h_hist <- max_annual_exotic_h %>% 
  ggplot(aes(x = max_exotic_h)) +
  geom_histogram()
#kind of normal...
exotic_h_qq <- max_annual_exotic_h %>% 
  ggplot(aes(sample = max_exotic_h)) +
  geom_qq()
#kind of normal...




### Column graph of H', by zone, by time since
exotic_h_col <- max_annual_exotic_h %>% 
  group_by(Time_Since, Replicate_Zone) %>% 
  summarize(mean = mean(max_exotic_h),
            se = std.error(max_exotic_h)) %>% 
  ggplot(aes(x = as.factor(Time_Since), y = mean)) +
  geom_col(fill = "coral3") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = .2)) +
  labs(title = "Exotic Species Diversity, over time", x = "Time Since Restoration (years)", y = "Shannon's Index of Diversity", caption = "(Upland: ANOVA F(7) = 16.68, p << 0.001; \n Tukey's sig 2-4+, sig 3-4+, 5-7, 5-9, 6-7 \n Central ANOVA F(7) = 5.264, p << 0.001; \n Tukey's sig 2-8, 3-8, 4-8, 5-8, 6-8, 7-8)") +
  facet_wrap(~Replicate_Zone) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme(text = element_text(size=20))
exotic_h_col

###ANOVA of H', for upland zone
exotic_h_u_aov <- max_annual_exotic_h %>% 
  filter(Replicate_Zone == "U") %>% 
  aov(max_exotic_h ~ as.factor(Time_Since), data = .)
#summary(exotic_h_u_aov)
#F(7) = 16.68, p << 0.001
####Tukey's Post-hoc
exotic_h_u_tukey <- TukeyHSD(exotic_h_u_aov)
#sig 2-4+, sig 3-4+, 5-7, 5-9, 6-7
###ANOVA of H', for transition zone
exotic_h_t_aov <- max_annual_exotic_h %>% 
  filter(Replicate_Zone == "T") %>% 
  aov(max_exotic_h ~ as.factor(Time_Since), data = .)
#summary(exotic_h_t_aov)
#F(7) = 3.496, p = 0.00138
####Tukey's Post-hoc
exotic_h_t_tukey <- TukeyHSD(exotic_h_t_aov)
#no sig
###ANOVA of H', for central zone
exotic_h_c_aov <- max_annual_exotic_h %>% 
  filter(Replicate_Zone == "C") %>% 
  aov(max_exotic_h ~ as.factor(Time_Since), data = .)
#summary(exotic_h_c_aov)
#F(7) = 5.264, p << 0.001
####Tukey's Post-hoc
exotic_h_c_tukey <- TukeyHSD(exotic_h_c_aov)
#sig 2-8, 3-8, 4-8, 5-8, 6-8, 7-8


## Boxplots exotic H', by zone, by time since
exotic_h_box <- max_annual_exotic_h %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "Central",  Replicate_Zone == "T" ~ "Transition", Replicate_Zone == "U" ~ "Upland")) %>%
  ggplot(aes(x = as.factor(Time_Since), y = max_exotic_h)) +
  geom_boxplot(fill = "coral3") +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Exotic Species Diversity, over time", x = "Age of Pool (years)", y = "Shannon's Index", caption = "Sig increase at 7 in central \n Increase in transition around 6 \n Increase in upland at 4 and 7 and 9 \n (Sig increase in central and upland over time)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 2))
exotic_h_box

###LMER
exotic_h_lmer <- lmer(max_exotic_h ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|plot_replicate), max_annual_exotic_h)
#summary(exotic_h_lmer)
#some age, age:upland and age:transition sig
#### Check diagnostic graphs
#E1 <- resid(exotic_h_lmer)
#F1 <- fitted(exotic_h_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_h_lmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_h_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_h_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_h_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal
#### ANOVA of lmer
exotic_h_lmer_anova <- anova(exotic_h_lmer)
#### Tukey's post-hoc of least-squares means
exotic_h_lmer_tukey <- emmeans(exotic_h_lmer, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(exotic_h_lmer_tukey)
#Central: 6-7 p = 0.0592
#Transition: 2-3 p = 0.0047, 3-4 p = 0.0085, 3-6 p = 0.01039, 3-7 p = 0.0454, 5-6 p = 0.0371, 5-7 p = 0.0059
#Upland: 2-7 p = 0.0038, 2-8 p = 0.1972, 2-9 p = 0.0085, 2-9 p << 0.001, 3-4 p << 0.001, 3-5 p - 0.0128, 3-7 p << 0.001, 3-8 p = 0.0044, 3-9 p = 0.001, 4-7 p = 0.014, 4-9 p = 0.0074, 5-7 p << 0.001, 5-8 p = 0.0613, 5-9 p << 0.001, 6-7 p << 0.001, 6-8 p = 0.0010, 6-9 p << 0.001, 8-9 p = 0.0430
exotic_h_continuous_lmer <- max_annual_exotic_h %>% 
  filter(Replicate_Zone == "U") %>% 
  lmer((max_exotic_h) ~ (Time_Since) + (1|Year) + (1|plot_replicate), data = .)
#summary(exotic_h_continuous_lmer)
#central and upland sig
#### Check diagnostic graphs
#E1 <- resid(exotic_h_continuous_lmer)
#F1 <- fitted(exotic_h_continuous_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(exotic_h_continuous_lmer)$plot_replicate[,1], main = "Quadrat Residuals")
#qqline(ranef(exotic_h_continuous_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(exotic_h_continuous_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(exotic_h_continuous_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal

```

## 1b. Do pools converge on the same community composition over time?

Hypothesis: Pools that are restored in the same way will converge on the same community over time.

Results: Some plots are sig different throughout (PH1, RT); Pools aged 2-6 are the same; pools aged 7-9 are the same


```{r echo = FALSE, message = FALSE, warning = FALSE}
# 1b. Do pools converge on the same community composition over time? (NMDS, cluster)

## Community matrix
np_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:141)

## NMDS, by plot
np_nmds <- metaMDS(np_community_matrix, trymax = 100, maxit = 999)
#stressplot(np_nmds)
#Generally, an average stress of > 0.2 is an indication that the ordination didn’t do a good job of representing community structure. However, NMDS stress increases as sample size increases, so if you have many samples (as this dataset does) your ordination will likely exhibit a lot of stress
#stress = .1843

np_nmds_scores <- scores(np_nmds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>% 
  full_join(np_community_matrix_variables)

np_nmds_fit <- envfit(np_nmds, np_community_matrix, perm = 999) #envfit() takes the output of metaMDS() and the species matrix you created

np_nmds_fit_pvals <- np_nmds_fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") #extract p-values for each species

np_nmds_fit_coords <- np_nmds_fit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., np_nmds_fit_pvals, by = "species") %>% 
  filter(pvals <= 0.001) #extract coordinates for species, only keep species with p-val <= 0.0001

### NMDS biplot, by plot
np_nmds_plot_species_biplot <- np_nmds_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = Plot)) +
  stat_ellipse(aes(color = Plot)) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species)) +
  theme_classic() +
  labs(title = "NMDS, by plot", caption = "PH1 has different community (post-hoc pairwise perMANOVA p = 0.021) \n RT also has slightly different community (post-hoc pairwise perMANOVA p = 0.021)")
np_nmds_plot_species_biplot

####perMANOVA, by plot
np_plot_permanova <- adonis(np_community_matrix ~ Plot, data = np_community_matrix_variables)
#F(6) = 7.6089, p = 0.001
####Post-hoc test
np_plot_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$Plot)
#PH1-RT p = .021, PH1-TP p = .021, PH1-MS p = .021, PH1-WT1 p = .021, PH1-WT2 p = .021, PH1-CS p = .021, RT-MS p = .021, RT-WT2 p = .021, RT-CS p = .021, TP-WT2 p = .063
#only non-sig are RT-TP, TP-MS, TP-WT1, TP-CS, MS-WT1, WT1-WT2, WT1-CS


### NMDS, by age, with species biplot
np_nmds_age_species_biplot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = as.factor(age))) +
  stat_ellipse(aes(color = as.factor(age))) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS, by pool age", caption = "Sig difference in community across age (perMANOVA p = 0.001) -- community changes at year 6")
np_nmds_age_species_biplot

####perMANOVA, by age
np_age_permanova <- adonis(np_community_matrix ~ age, data = np_community_matrix_variables)
#F(6) = 15.499, p = 0.001
####Post-hoc test
np_age_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$age)
#7-6 p = .028, 2-6 p = .028, 7-5 p = .028, 7-4 p = .056, 7-3 p = .028, 7-2 p = .028, 8-6 p = .028, 8-5 p = .028, 8-4 p = .028, 8-3 p = .028, 8-2 p = .028, 9-2 p = .028, 9-3 p = .028, 9-4 p = .056, 9-5 p = .028, 9-6 p = .028


### NMDS
np_nmds_zone <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = Replicate_Zone)) +
  stat_ellipse(aes(color = Replicate_Zone)) +
  theme_classic() +
  labs(title = "NMDS, by zone", caption = "Central zone is sig different from  transition and upland zones (perMANOVA p = 0.067)") +
  scale_color_manual(values = c("dodgerblue4", "darkolivegreen3", "goldenrod"), label = c("Central", "Transition", "Upland"), name = "Zone")
np_nmds_zone

####perMANOVA, by zone
np_zone_permanova <- adonis(np_community_matrix ~ Replicate_Zone, data = np_community_matrix_variables)
#F(2) = 1.6991, p = 0.062
####Post-hoc test
np_zone_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$Replicate_Zone)
#C-T p = 0.123, C-U p = 0.096, T-U p = 1.000

####Partition-based cluster analysis plot, for zone
np_kmeans <- kmeans(np_nmds_scores[2:3], 3)

np_clusters <- data.frame(np_nmds_scores, cluster_no = factor(np_kmeans$cluster))

np_clusters_plot <- ggplot(np_clusters) +
  geom_point(aes(x = NMDS1, y = NMDS2, color = cluster_no, pch = Replicate_Zone)) +
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Replicate_Zone))
  labs(title = "Partition-based cluster analysis, by zone")
np_clusters_plot

```

## 1c. How does the increase in exotics affect the community?

Hypothesis: Exotic annual grass species will consistently dominate the community and outcompete native species.

Results: Native cover remains high over time, suggesting coexistence

```{r echo = FALSE, message = FALSE, warning = FALSE}
# 1c. How does the increase in exotics affect the community? (column graphs, NMDS)

## Calculate total native % cover per quadrat, find max total native % cover per quadrat per year, add time since restoration column
np_annual_total_native <- np_percent_cover %>%
  filter(Native_Status == "N") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_native_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  group_by(Year, Plot, Replicate, Replicate_Zone) %>% 
  summarize(max_annual_native_pc = max(total_native_pc)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  mutate(plot_replicate = paste(Plot, Replicate, sep = "-"))

### Check for normality
native_pc_hist <- ggplot(np_annual_total_native, aes(x = max_annual_native_pc)) +
  geom_histogram()
#pretty normal
native_pc_qq <- ggplot(np_annual_total_native, aes(sample = max_annual_native_pc)) +
  geom_qq()
#pretty normal


## Column graph for native cover, by zone, by time since
native_pc_col <- np_annual_total_native %>% 
  group_by(Replicate_Zone, Time_Since) %>% 
  summarize(mean = mean(max_annual_native_pc),
            se = std.error(max_annual_native_pc)) %>% 
  ggplot(aes(x = as.factor(Time_Since))) +
  geom_col(aes(y = mean), fill = "darkgreen") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = .2)) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Total Native  Percent Cover, over time", x = "Age of Pool (years)", y = "Percent Cover (%)", caption = "Increase in native cover at year 3 in central and upland zones \n (Sig increase in central over time)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 170))
native_pc_col


#### ANOVA of native cover
native_pc_anova <- aov(max_annual_native_pc ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|as.complex(plot_replicate)), data = np_annual_total_native)
#summary(native_pc_anova)
#age, zone sig
#### Tukey's post-hoc
native_pc_tukey <- emmeans(native_pc_anova, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(native_pc_tukey, comparisons = TRUE)
#Central: 2-3 p = 0.0022, 2-4 p = 0.0001, 2-5 p << 0.001, 2-6 p = 0.0022, 2-7 p =.0013, 2-8 p = 0.0006, 2-9 p << 0.001
#Transition: no sig
#Upland: 2-4 p = 0.1440, 2-5 p = 0.0419, 2-6 p = 0.1973, 2-8 p = 0.1732, 2-9 p = 0.0094
#### LMER of native cover (continuous)
native_pc_continuous_lmer <- np_annual_total_native %>% 
  filter(Replicate_Zone == "C") %>% 
  lmer((max_annual_native_pc) ~ (Time_Since) + (1|Year) + (1|plot_replicate), data = .)
#summary(native_pc_continuous_lmer)
#only central sig
#### Check diagnostic graphs
#E1 <- resid(native_pc_continuous_lmer)
#F1 <- fitted(native_pc_continuous_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(native_pc_continuous_lmer)$plot_replicate[,1], main = "Pool Residuals")
#qqline(ranef(native_pc_continuous_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(native_pc_continuous_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(native_pc_continuous_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal




## Shannon's Index for native community of each plot per year
native_h <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "N") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  group_by(date, month_number, Year, Plot, Replicate, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(8:69) %>% 
  diversity() %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>%
  rename("h" = ".")

## Max H' for each quadrat per zone per year, add time since column
max_annual_native_h <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "N") %>% 
  group_by(date, month_number, Year, Plot, Replicate, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  select(1:8) %>% 
  full_join(native_h) %>% 
  group_by(Year, Plot, Replicate, Replicate_Zone) %>% 
  mutate(max_native_h = max(h)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  mutate(plot_replicate = paste(Plot, Replicate, sep = "-"))

### Exploratory graphs
native_h_hist <- max_annual_native_h %>% 
  ggplot(aes(x = max_native_h)) +
  geom_histogram()
#not normal
native_h_qq <- max_annual_native_h %>% 
  ggplot(aes(sample = max_native_h)) +
  geom_qq()
#not normal

## Boxplot  for native H', by zone, by time since
native_h_time_box <- max_annual_native_h %>% 
  ggplot(aes(x = as.factor(Time_Since), y = max_native_h)) +
  geom_boxplot(fill = "darkgreen") +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Native Shannon's Index, over time", x = "Age of Pool (years)", y = "Shannon's Index", caption = "Decrease in 3 in central and upland, \n increase around 4, 7, 9 in central and upland \n increase 4, 9 in transition \n (Sig increase in central and upland over time)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 3))
native_h_time_box

### LMER for native H'
native_h_lmer <- lmer((max_native_h) ~ as.factor(Time_Since) + Replicate_Zone + as.factor(Time_Since):Replicate_Zone + (1|Year) + (1|plot_replicate), max_annual_native_h)
#summary(native_h_lmer)
#some age, zone, age:zone sig
#### Check diagnostic graphs
#E1 <- resid(native_h_lmer)
#F1 <- fitted(native_h_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance for Native H'") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(native_h_lmer)$plot_replicate[,1], main = "Pool Residuals")
#qqline(ranef(native_h_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(native_h_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(native_h_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal
#### ANOVA of lmer
native_h_lmer_anova <- anova(native_h_lmer)
#### Tukey's post-hoc of least-squares means
native_h_lmer_lmer_tukey <- emmeans(native_h_lmer, pairwise ~ as.factor(Time_Since) | Replicate_Zone, adjust = "tukey")
#plot(native_h_lmer_lmer_tukey)
#Central: 2-3 p = 0.0006, 2-4 p = 0.0605, 2-9 p = 0.0642, 3-6 p << 0.001, 3-6 p = 0.0977, 3-7 p = 0.022, 3-8 p = 0.0003, 3-9 p = 0.0002, 4-5 p << 0.001, 4-7 p = 0.0016, 4-8 p = 0.0001, 4-9 p = 0.0001, 5-8 p = 0.0132, 5-9 p = 0.0012, 6-7 p = 0.0061, 6-8 p << 0.001, 6-9 p << 0.001, 7-8 p = 0.0377, 7-9 p = 0.0003
#Transition: 4-7 p = 0.0753, 5-7 p = 0.0455, 7-8 p = 0.0056, 8-9 p = 0.0097
#Upland: 2-3 p = 0.0569, 2-4 p = 0.0497, 2-7 p = 0.0255, 3-8 p = 0.1933, 3-9 p = 0.1414, 3-4 p << 0.001, 3-5 p = 0.002, 3-6 p = 0.0479, 3-7 p = 0.0004, 3-8 p = 0.0072, 3-9 p = 0.0110, 4-7 p = 0.0705, 5-7 p = 0.0001, 5-8 p = 0.1666, 5-9 p = 0.0876, 6-7 p << 0.001, 6-8 p = 0.0216, 6-9 p = 0.0127
native_h_continuous_lmer <- max_annual_native_h %>% 
  filter(Replicate_Zone == "T") %>% 
  lmer((max_native_h) ~ (Time_Since) + (1|Year) + (1|plot_replicate), data = .)
#summary(native_h_continuous_lmer)
#central, upland slightly sig
#### Check diagnostic graphs
#E1 <- resid(native_h_continuous_lmer)
#F1 <- fitted(native_h_continuous_lmer)
#plot(x=F1, y=E1, xlab="Fitted Model Values", ylab="Pearson Residuals", main="Variance for Native H'") #run 2 lines together -- assessing homoskedasticity -- you want to see no pattern
#abline(h=0)
#qqnorm(E1, ylab="Pearson Residuals")
#qqline(E1) #run 2 lines together -- assess if model residuals are normally distributed
#qqnorm(ranef(native_h_continuous_lmer)$plot_replicate[,1], main = "Pool Residuals")
#qqline(ranef(native_h_continuous_lmer)$plot_replicate[,1]) #run 2 lines together -- assess if random effects residuals are normal
#qqnorm(ranef(native_h_continuous_lmer)$Year[,1], main = "Monitoring Year Residuals")
#qqline(ranef(native_h_continuous_lmer)$Year[,1]) #run 2 lines together -- assess if random effects residuals are normal







## Dominant natives/nonnatives
np_dominant <- np_percent_cover %>%
  drop_na(percent_cover) %>% 
  group_by(Species_Full_Name) %>% 
  summarize(max_pc = max(percent_cover)) %>% 
  full_join(metadata) %>% 
  drop_na(max_pc) %>% 
  filter(Native_Status != "NA") %>% 
  filter(max_pc >=10)
#dominant exotics: FEPE, FEMY, POMO, CETE, BRDI, GEDI, PLCO, POAN, SPSP, BRHO, CRSC, LYHY
#dominant natives: ELMA, STPU, JUPH, JUME, HOBR, BOMA, CAPR, CYER, JUBU, ERVA, ALSA, JUPA, MALE, SIBE, CEPA, ELTR, EPCA, ELGL, SCPU, SYSU, DEFA, FRSA

## Average natives/nonnatives per zone
np_zone_richness <- np_percent_cover %>%
  filter(Native_Status != "NA") %>% 
  filter(Year > 2016) %>% 
  group_by(Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(mean = mean(percent_cover)) %>% 
  filter(mean>0) %>% 
  drop_na() %>% 
  group_by(Plot, Replicate_Zone, Native_Status) %>% 
  summarize(richness = length(mean)) %>% 
  group_by(Replicate_Zone, Native_Status) %>% 
  summarize(mean_richness = mean(richness))





## Bray-Curtis Dissimilarity Index

### Bray-Curtis of all species
np_bray <- as.matrix(vegdist(np_community_matrix, method = "bray"))
#hist(np_bray, xlab = "ecological distance", main = "np_bray")

### Exotic community matrix, central
exotic_c_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "E") %>%
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  filter(Replicate_Zone == "1") %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:27)

#### Bray-Curtis of exotics, central
np_exotic_c_bray <- as.matrix(vegdist(exotic_c_community_matrix, method = "bray")) %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "quadrat_id", values_to = "bray_curtis", 1:59) %>%
  add_column(replicate_zone = "Central") %>% 
  add_column(native_status = "E")
#hist(np_exotic_c_bray, xlab = "ecological distance", main = "np_exotic_c_bray")

### Exotic community matrix, transition
exotic_t_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "E") %>%
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  filter(Replicate_Zone == "2") %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:57)

#### Bray-Curtis of exotics, transition
np_exotic_t_bray <- as.matrix(vegdist(exotic_t_community_matrix, method = "bray")) %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "quadrat_id", values_to = "bray_curtis", 1:127) %>%
  add_column(replicate_zone = "Transition") %>% 
  add_column(native_status = "E")
#hist(np_exotic_t_bray, xlab = "ecological distance", main = "np_exotic_t_bray")

### Exotic community matrix, upland
exotic_u_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "E") %>%
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  filter(Replicate_Zone == "3") %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:62)

#### Bray-Curtis of exotics, upland
np_exotic_u_bray <- as.matrix(vegdist(exotic_u_community_matrix, method = "bray")) %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "quadrat_id", values_to = "bray_curtis", 1:135) %>%
  add_column(replicate_zone = "Upland") %>% 
  add_column(native_status = "E")
#hist(np_exotic_u_bray, xlab = "ecological distance", main = "np_exotic_u_bray")

### Native community matrix, central
native_c_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "N") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  filter(Replicate_Zone == "1") %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  drop_na(Replicate_Zone) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:34)

#### Bray-Curtis of natives, central
np_native_c_bray <- as.matrix(vegdist(native_c_community_matrix, method = "bray")) %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "quadrat_id", values_to = "bray_curtis", 1:132) %>%
  add_column(replicate_zone = "Central") %>% 
  add_column(native_status = "N")
#hist(np_native_c_bray, xlab = "ecological distance", main = "np_native_c_bray")

### Native community matrix, transition
native_t_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "N") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  filter(Replicate_Zone == "2") %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  drop_na(Replicate_Zone) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:53)

#### Bray-Curtis of natives, transition
np_native_t_bray <- as.matrix(vegdist(native_t_community_matrix, method = "bray")) %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "quadrat_id", values_to = "bray_curtis", 1:182) %>%
  add_column(replicate_zone = "Transition") %>% 
  add_column(native_status = "N")
#hist(np_native_t_bray, xlab = "ecological distance", main = "np_native_t_bray")

### Native community matrix, upland
native_u_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "N") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  filter(Replicate_Zone == "3") %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  drop_na(Replicate_Zone) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:44)

#### Bray-Curtis of natives, upland
np_native_u_bray <- as.matrix(vegdist(native_u_community_matrix, method = "bray")) %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "quadrat_id", values_to = "bray_curtis", 1:188) %>%
  add_column(replicate_zone = "Upland") %>% 
  add_column(native_status = "N")
##hist(np_native_u_bray, xlab = "ecological distance", main = "np_native_u_bray")


#### Dataframe of all Bray-Curtis
bray_df <- rbind(np_exotic_c_bray, np_exotic_t_bray, np_exotic_u_bray, np_native_c_bray,  np_native_t_bray, np_native_u_bray) %>% 
  mutate(status_zone = paste(native_status, replicate_zone, sep = "_"))

### Boxplots of Bray-Curtis
bray_box <- bray_df %>% 
  ggplot() +
  geom_boxplot(aes(x = replicate_zone, y = bray_curtis, fill = native_status)) +
  labs(title = "Bray-Curtis Dissimilarity", x = "Vegetation Zone", y = "Bray-Curtis Distance", caption  = "Kruskal-Wallis p << 0.001") +
  theme_classic() +
  scale_fill_manual(values = c("coral3", "darkgreen"), label = c("Exotic", "Native"), name = "Native Status")
bray_box

##### Kruskal-Wallis
bray_kw <- kruskal.test(bray_curtis ~ as.factor(status_zone), data = bray_df)
#p << .001
### Post-hoc Dunn's test
bray_dunn <- dunnTest(bray_curtis ~ as.factor(status_zone), data = bray_df)
#all sig diff

```