---
title: "NP Draft 1: Aggregate all pools together"
author: "Joanna Tang"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

# Loading packages, read in data
library(tidyverse)
library(knitr)
library(kableExtra)
library(vegan)
library(car)
library(RColorBrewer)
library(reshape2)
library(knitr)
library(kableExtra)
np_2016_seedbank_master <- read_csv("np_2016_seedbank.csv")
np_percent_cover_master <- read_csv("np_percent_cover.csv")
metadata <- read_csv("metadata.csv")

```

```{r include = FALSE}

# Tidy up data

## Separate "Replicate" label into Replicate_Zone and Replicate_Number, join with species metadata, convert month and year to "date" column
np_percent_cover <- np_percent_cover_master %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3"))
np_percent_cover <- full_join(np_percent_cover, metadata) %>% 
  mutate(month_number = case_when(Month == "JANUARY" ~ 1, Month == "FEBRUARY" ~ 2, Month == "MARCH" ~ 3, Month == "APRIL" ~ 4, Month == "MAY" ~ 5, Month == "JUNE" ~ 6, Month == "JULY" ~ 7, Month == "AUGUST" ~ 8, Month == "SEPTEMBER" ~ 9, Month == "OCTOBER" ~ 10, Month == "NOVEMBER" ~ 11, Month == "DECEMBER" ~ 12)) %>% 
  mutate(date = paste(Year, month_number, "1", sep="-" ))
np_percent_cover$date = as.Date(np_percent_cover$date)

# Convert np_2016_seedbank into tidy format, remove NAs, separate "Replicate" label into Replicate_Zone and Replicate_Number
seedbank_2016 <- gather(np_2016_seedbank_master, key = "Species", value = "Count", -c(Plot, Replicate)) %>% 
  na.omit(np_2016_seedbank_master) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3")) %>% 
  full_join(metadata)


```


```{r include = FALSE}

# Outline
# 1. How does exotic species abundance and diversity correlate with time since restoration?
# 1a. compare max % cover of each exotic species, 2017 vs 2018
# 1b. compare max % cover of each exotic species, 2017 vs 2018, by zone

#2. Do the invading exotic species exhibit a similar set of traits?
# 2a. annual vs. perennial graphs of c to to t u
# 2b. max % cover vs. % in seed bank
# 2c. phenology
# 2d. moisture regime

```

# 1. How does exotic species percent cover and diversity correlate with time since restoration?

Hypothesis: We will see increased exotic species percent cover and diversity with increased time since restoration.

# 1a. Total percent cover over time since restoration

Results: Increase in total exotic percent cover over time.

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1a. Compare total % cover of exotics (sum of max % cover of each species) over time since restoration


## Create new data frame with creation year of each pool
creation_year_df <- data.frame(Plot = c("PH1", "TP", "RT", "MS", "WT1", "WT2", "CS"), Creation_Year = c("2010", "2012", "2012", "2015", "2014", "2014", "2013"))

## Add time since restoration column
np_percent_cover_time_cover <- np_percent_cover %>%
  filter(Native_Status == "E", Year >2016) %>% 
  group_by(Species_Full_Name, Year, Plot) %>% 
  summarize(max = max(percent_cover)) %>% 
  group_by(Year, Plot) %>% 
  summarize(total = sum(max)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year)
  
# Scatter plot of total % cover vs. time since
max_exotic_time_scatter <- np_percent_cover_time_cover %>% 
  ggplot(aes(x = as.factor(Time_Since), y = total)) +
  geom_point(aes(color = Plot)) +
  geom_line(aes(color = Plot)) +
  geom_smooth(method = lm, aes(x = Time_Since), se = TRUE, size = .5) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration", y = "Percent Cover", caption = "More variation as time increases; PH1, CS, TP decreased")
max_exotic_time_scatter


# Boxplot of total % cover vs. time since
max_exotic_time_box <- np_percent_cover_time_cover %>% 
  ggplot(aes(x = as.factor(Time_Since), y = total)) +
  geom_boxplot(width = .5) +
  geom_jitter(aes(color = Plot), width = .1) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration", y = "Percent Cover", caption = "More variation as time increases; PH1, CS, TP decreased")
max_exotic_time_box

```


## 1b. Species richness over time since restoration

Results: Increase in species richness over time

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1a. Compare species richness over time since restoration

## Add time since restoration column
np_percent_cover_time_richness <- np_percent_cover %>%
  filter(Native_Status == "E", Year >2016) %>% 
  group_by(Species_Full_Name, Year, Plot) %>% 
  summarize(max = max(percent_cover)) %>% 
  group_by(Year, Plot) %>% 
  summarize(richness = length(max)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year)
  

## Scatter plot of species richness vs. time since
richness_time_scatter <- np_percent_cover_time_richness %>% 
  ggplot(aes(x = as.factor(Time_Since), y = richness)) +
  geom_point(aes(color = Plot)) +
  geom_line(aes(color = Plot)) +
  geom_smooth(method = lm, aes(x = Time_Since), se = TRUE, size = .5) +
  labs(title = "Species Richness of Exotics", x = "Time Since Restoration", y = "Species Richness", caption = "More variation as time increases; PH1, CS, MS decreased")
richness_time_scatter


# Boxplot of species richness vs. time since
richness_time_box <- np_percent_cover_time_richness %>% 
  ggplot(aes(x = as.factor(Time_Since), y = richness)) +
  geom_boxplot(width = .5) +
  geom_jitter(aes(color = Plot), width = .1) +
  labs(title = "Species Richness of Exotics", x = "Time Since Restoration", y = "Species Richness")
richness_time_box

```

## 1c. Grouped by zone

Results: Transition and Upland zones increased in exotic species by the most.

```{r echo = FALSE, warning = FALSE, message = FALSE}

# 1c. Compare total % cover of exotic species (sum of max % cover of each species) and species richness over time since by zone

## Column graph of total % cover vs. time since by zone
max_exotic_time_zone_col <- np_percent_cover %>%
  filter(Native_Status == "E", Year >2016) %>% 
  group_by(Species_Full_Name, Year, Plot, Replicate_Zone) %>% 
  summarize(max = max(percent_cover)) %>% 
  group_by(Year, Plot, Replicate_Zone) %>% 
  summarize(total = sum(max),
            richness = length(max)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  ggplot(aes(x = Replicate_Zone)) +
  geom_col(aes(y = total, fill = as.factor(Time_Since)), position = "dodge") +
  labs(title = "Total Percent Cover of Exotics", x = "Zone", y = "Percent Cover", caption = "Transition and Upland zones increased much more")
max_exotic_time_zone_col


## Column graph of species richness vs. time since by zone
richness_time_zone_col <- np_percent_cover %>%
  filter(Native_Status == "E", Year >2016) %>% 
  group_by(Species_Full_Name, Year, Plot, Replicate_Zone) %>% 
  summarize(max = max(percent_cover)) %>% 
  group_by(Year, Plot, Replicate_Zone) %>% 
  summarize(total = sum(max),
            richness = length(max)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  ggplot(aes(x = Replicate_Zone)) +
  geom_col(aes(y = richness, fill = as.factor(Time_Since)), position = "dodge") +
  labs(title = "Species Richness of Exotics", x = "Zone", y = "Species Richness", caption = "Transition zone increased most")
richness_time_zone_col

```


# 2. Do the invading exotic species exhibit a similar set of traits?

## 2a. 2016 Seed Bank vs. 2017 Percent Cover

```{r echo = FALSE, warning = FALSE, message = FALSE}

#2a. Compare % in 2016 seed bank vs. max % cover in 2017 of each exotic species, grouped by time since restoraitn

## Mean % count of each exotic species
seed_bank_summary <- seedbank_2016 %>%
  filter(Native_Status == "E") %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = 2017 - Creation_Year) %>% 
  group_by(Species_Full_Name, Plot, Time_Since) %>% 
  summarize(mean = mean(Count))

## Scatter plot of mean count in 2016 seed bank vs. max % cover in 2017 of each exotic species, colored by time since restoration
seed_bank_cover_scatter <- np_percent_cover %>%
  filter(Native_Status == "E", Year == 2017) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = 2017 - Creation_Year) %>% 
  group_by(Species_Full_Name, Plot, Time_Since) %>% 
  summarize(max = max(percent_cover)) %>% 
  group_by(Plot, Species_Full_Name, Time_Since) %>% 
  summarize(total = sum(max)) %>% 
  full_join(seed_bank_summary) %>% 
  filter(Plot != "NA") %>% 
  mutate_all(funs(replace(., is.na(.), 0))) %>% 
  ggplot(aes(x = mean, y = total)) +
  geom_point(aes(color = Time_Since)) +
  labs(title = "2016 Exotic Seed Bank vs. 2017 Exotic Percent Cover", x = "Abundance in seed bank", y = "Percent Cover")
seed_bank_cover_scatter


```

## 2b. Max exotic percent cover vs. pool depth

```{r}

depth_df <- data.frame(depth = c("12", "13", "13", "12", "17.75", "13", "15.5", "12", "13", "13", "12", "17.75", "13", "15.5"), Plot = c("CS", "MS", "PH1", "RT", "TP", "WT1", "WT2", "CS", "MS", "PH1", "RT", "TP", "WT1", "WT2"), total = np_percent_cover_depth$total)

np_percent_cover_depth <- np_percent_cover %>%
  filter(Native_Status == "E", Year >2016) %>% 
  group_by(Species_Full_Name, Year, Plot) %>% 
  summarize(max = max(percent_cover)) %>% 
  group_by(Year, Plot) %>% 
  summarize(total = sum(max)) %>% 
  full_join(depth_df)
  
## Scatter plot of total % cover vs. time since
depth_exotic_scatter <- depth_df %>% 
  ggplot(aes(x = as.numeric(depth), y = total, group = Plot)) +
  geom_point(aes(color = Plot)) +
  geom_line(aes(color = Plot)) +
  geom_smooth(method = lm, aes(group = NULL), size = .2) +
  labs(title = "Total Percent Cover of Exotics", x = "Depth of Pool", y = "Percent Cover", caption = "Deeper pools have less exotics")
depth_exotic_scatter

```

