---
title: "NP Draft 5"
author: "Joanna Tang"
date: "9/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ESA Talk

```{r include = FALSE}

# Loading packages, read in data
library(tidyverse)
library(knitr)
library(kableExtra)
library(vegan)
library(car)
library(RColorBrewer)
library(reshape2)
library(knitr)
library(kableExtra)
library(FSA)
library(plotly)
library(ggsignif)
library(lme4)
library(bestNormalize)
library(betareg)
library(nlme)
library(MASS)
library(fitdistrplus)
library(plotrix)

np_percent_cover_master <- read_csv("np_percent_cover.csv")
metadata <- read_csv("metadata.csv")

```


```{r include = FALSE}

# Outline

# 0. Tidy up data

# 1. How does exotic community change over time?
## 1a. How does total exotic percent cover change with time since restoration (and interannual climate variability?)? (boxplots)
## 1b. Is there a threshold (# of years) after which exotic abundance increases? (linear model)
## 1c. Do pools converge on the same exotic community composition over time? (NMDS)

# 2. What environmental filters can be manipulated to decrease exotics?
## 2a. How does the seedbank compare to the aboveground cover? (presence/absence data)
## 2b. Does active management result in a different exotic community? (NMDS w/ old Del Sol data)
## 2c. Do certain exotics and natives occupy the same (phenological) niche? (NMDS, or scatterplot b/t 2 species)

# 3. What is the ideal monitoring plan?
## 3a. Same as 1b -- is there a threshold after which exotics increase (i.e., what's the ideal duration of monitoring?)?
## 3b. What is the ideal frequency/season of monitoring? (boxplots)


```


```{r include = FALSE}

# 0. Tidy up data

## Separate "Replicate" label into Replicate_Zone and Replicate_Number, join with species metadata, convert month and year to "date" column
np_percent_cover <- np_percent_cover_master %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3"))
np_percent_cover <- full_join(np_percent_cover, metadata) %>% 
  mutate(month_number = case_when(Month == "JANUARY" ~ 1, Month == "FEBRUARY" ~ 2, Month == "MARCH" ~ 3, Month == "APRIL" ~ 4, Month == "MAY" ~ 5, Month == "JUNE" ~ 6, Month == "JULY" ~ 7, Month == "AUGUST" ~ 8, Month == "SEPTEMBER" ~ 9, Month == "OCTOBER" ~ 10, Month == "NOVEMBER" ~ 11, Month == "DECEMBER" ~ 12)) %>% 
  mutate(date = paste(Year, month_number, "1", sep="-" ))
np_percent_cover$date = as.Date(np_percent_cover$date)

```

# 1. How does exotic community change over time?

## 1a. How does total exotic percent cover change with time since restoration?

Hypothesis: Exotics will be able to reinvade, causing increased exotic abundance and diversity with increased time since restoration.

Results: Significant increase in total exotic percent cover after year 6 -- due to increase in exotics in transition and upland zones.

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1a. How does total exotic percent cover change with time since restoration (and interannual climate variability?)? (boxplots)


## Create new data frame with creation year of each pool
creation_year_df <- data.frame(Plot = c("PH1", "TP", "RT", "MS", "WT1", "WT2", "CS"), Creation_Year = c("2010", "2012", "2012", "2015", "2014", "2014", "2013"))

## Calculate total exotic % cover per quadrat, find max total exotic % cover per quadrat per year, add time since restoration column
np_annual_total_exotic <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  group_by(Year, Plot, Replicate, Replicate_Zone) %>% 
  summarize(max_annual_exotic_pc = max(total_exotic_pc)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year)

## Dataframe of median of the max annual total % cover of each pool
np_annual_total_exotic_medians <- np_annual_total_exotic %>% 
  group_by(Year, Plot, Time_Since) %>% 
  summarize(median = median(max_annual_exotic_pc))

## Boxplot of total exotic % cover
total_exotic_pc_box <- np_annual_total_exotic %>% 
  ggplot(aes(x = as.factor(Time_Since), y = max_annual_exotic_pc)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Increase in median after year 6 driven by sig incraese in transition and upland zones (Dunn's Kruskal-Wallis Multiple Comparisons p =)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
total_exotic_pc_box

## Boxplot of total exotic % cover, by zone
total_exotic_pc_zone_box <- np_annual_total_exotic %>% 
  ggplot(aes(x = as.factor(Time_Since), y = max_annual_exotic_pc)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Total Percent Cover of Exotics, by zone", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Increase in median after year 6 driven by sig incraese in transition and upland zones (Dunn's Kruskal-Wallis Multiple Comparisons p =)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
total_exotic_pc_zone_box

## Boxplot of total exotic % cover, by monitoring year -- everything's higher in 2019...
total_exotic_pc_year_box <- np_annual_total_exotic %>% 
  ggplot(aes(x = as.factor(Year), y = max_annual_exotic_pc)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  labs(title = "Total Percent Cover of Exotics", x = "Year", y = "Percent Cover", caption = "everything's higher in 2019...") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
total_exotic_pc_year_box




# Check for normality -- not normal
pc_hist <- ggplot(np_annual_total_exotic, aes(x = max_annual_exotic_pc)) +
  geom_histogram()
pc_qq <- ggplot(np_annual_total_exotic, aes(sample = max_annual_exotic_pc)) +
  geom_qq()

#### from original data analysis for Cal-IPC poster
## 95% CI around median
#lower CI = (n/2) - (1.96*sqrt(n))/2 th ranked value
#upper CI = 1 + (n/2) + (1.96*sqrt(n))/2 th ranked value

#n(3yrs) = 3
#lower CI = (3/2) - (1.96*sqrt(3/2)) = -.9 -> 1st ranked value = 2.38
#upper CI = 1 + (3/2) + (1.96*sqrt(3/2)) = 4.9005 -> 3rd ranked value -> 3.0

#n(4yrs) = 4
#lower CI = (4/2) - (1.96*sqrt(4/2)) = -.77 -> 1st ranked value = 2.73
#upper CI = 1 + (4/2) + (1.96*sqrt(4/2)) = 5.77 -> 4th ranked value -> 4.27

#n(5yrs) = 5
#lower CI = (5/2) - (1.96*sqrt(5/2)) = -.60 -> 1st ranked value = 2.65
#upper CI = 1 + (5/2) + (1.96*sqrt(5/2)) = 6.60 -> 5th ranked value -> 11.533

#n(6yrs) = 3
#lower CI = (3/2) - (1.96*sqrt(3/2)) = -.9 -> 1st ranked value = 4.90
#upper CI = 1 + (3/2) + (1.96*sqrt(3/2)) = 4.9005 -> 3rd ranked value -> 17.86

#n(7yrs) = 3
#lower CI = (3/2) - (1.96*sqrt(3/2)) = -.9 -> 1st ranked value = 8.60
#upper CI = 1 + (3/2) + (1.96*sqrt(3/2)) = 4.9005 -> 3rd ranked value -> 19.12

exotic_pc_ci_df <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  filter(Time_Since >2) %>% 
  filter(Time_Since <8) %>% 
  group_by(Time_Since) %>% 
  summarize(
    median = median(total_exotic_pc),
    max = max(total_exotic_pc),
    min = min(total_exotic_pc)
  )

exotic_pc_ci_zone_df <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  filter(Time_Since >2) %>% 
  filter(Time_Since <8) %>% 
  group_by(Time_Since, Replicate_Zone) %>% 
  summarize(
    median = median(total_exotic_pc),
    max = max(total_exotic_pc),
    min = min(total_exotic_pc)
  )

#### from original data analysis for Cal-IPC poster
#n(3yrs) = 23
#lower CI = (23/2) - (1.96*sqrt(23/2)) = 4.85 -> 4th ranked value = 1.33
#upper CI = 1 + (23/2) + (1.96*sqrt(23/2)) = 19.15 -> 20th ranked value -> 3.50

#n(4yrs) = 30
#lower CI = (30/2) - (1.96*sqrt(30/2)) = 7.41 -> 7th ranked value = 2.00
#upper CI = 1 + (30/2) + (1.96*sqrt(30/2)) = 23.59 ->24th ranked value -> 4.33

#n(5yrs) = 34
#lower CI = (34/2) - (1.96*sqrt(34/2)) = 9.92 -> 9th ranked value = 2.50
#upper CI = 1 + (34/2) + (1.96*sqrt(34/2)) = 26.08 -> 26th ranked value -> 8.67

#n(6yrs) = 23
#lower CI = (23/2) - (1.96*sqrt(23/2)) = 4.85 -> 4th ranked value = 2.00
#upper CI = 1 + (23/2) + (1.96*sqrt(23/2)) = 19.15 -> 20th ranked value -> 15.00

#n(7yrs) = 22
#lower CI = (22/2) - (1.96*sqrt(22/2)) = 4.50 -> 4th ranked value = 2.50
#upper CI = 1 + (22/2) + (1.96*sqrt(22/2)) = 18.50 -> 18th ranked value -> 25.33

## Column graph of total % cover vs. time since, w/ 95% CI
total_exotic_col <- exotic_pc_ci_df %>% 
  ggplot(aes(x = as.factor(Time_Since), y = median)) +
  geom_col() +
  geom_errorbar(aes(ymin = min, ymax = max, width = .2)) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Sig increase in median during year 6 (Dunn's Kruskal-Wallis Multiple Comparisons p = )") +
  theme_classic() +
    scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 25)) +
  theme(text = element_text(size=20))
#total_exotic_col

## Column graph of total % cover vs. time since, w/ 95% CI, by zone
total_avg_exotic_zone_col <- exotic_pc_ci_zone_df %>% 
   ggplot(aes(x = as.factor(Time_Since), y = median)) +
  geom_col() +
  geom_errorbar(aes(ymin = min, ymax = max, width = .2)) +
  facet_wrap(~Replicate_Zone) +
  geom_signif(comparisons = list(c("5", "6")), map_signif_level = TRUE) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Sig increase in median during year 6 driven by sig incraese in transition zone (Dunn's Kruskal-Wallis Multiple Comparisons p = )") +
  theme_classic() +
      scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 30)) +
  theme(text = element_text(size=20))
#total_avg_exotic_zone_col

```


## 1b. Is there a threshold (# of years) after which exotic abundance increases?

Hypothesis: We will see increased exotic species percent cover and diversity with increased time since restoration due to reinvasion.

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1b. Is there a threshold (# of years) after which exotic abundance increases? (linear model)

# Scatterplot of median of annual max total % exotic cover
np_annual_total_exotic_medians_scatter <- np_annual_total_exotic_medians %>% 
  ggplot(aes(x = Time_Since, y = median, group = Plot)) +
  geom_point(aes(color = Plot)) +
  geom_smooth(method = lm, aes(x = Time_Since, group = NULL), se = TRUE, size = .5) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Median Percent Cover (%)", caption = "High variation; WT2, PH1, & esp WT1 and RT increased") +
  theme_classic() +
  scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0))
np_annual_total_exotic_medians_scatter

```



## 1c. Do pools converge on the same exotic community composition over time?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1c. Do pools converge on the same exotic community composition over time? (NMDS)

# From np_draft3.Rmd -- Compare 2016 exotic seed bank species to May 2017 anad 2018 percent cover species, by pool and zone

## Take out zeros in 2016 seed bank data
seedbank_2016[seedbank_2016 == 0] <- NA

## Calculate exotic and native species richness for 2016 seed bank data
richness_2016_seed_bank <- seedbank_2016 %>% 
  add_column(Year = 2016) %>% 
  select(Plot, Replicate, Replicate_Zone, Year, Native_Status, Count) %>% 
    na.omit() %>%
  group_by(Plot, Replicate, Replicate_Zone, Year, Native_Status) %>% 
  summarize(richness = length(Count))

## Calculate exotic and native species richness for May 2017/2018 percent cover data, then join with 2016 seed bank data
np_seed_bank_cover_may <- np_percent_cover %>% 
  filter(month_number == 5) %>% 
  group_by(Plot, Replicate, Replicate_Zone, Year, Native_Status) %>% 
  summarize(richness = length(percent_cover)) %>% 
  full_join(richness_2016_seed_bank) %>% 
  na.omit()

## Calculate mean exotic species richness, by zone in each plot
np_exotic_richness_may <- np_seed_bank_cover_may %>% 
  filter(Native_Status == "E") %>% 
  group_by(Year, Plot, Replicate_Zone) %>% 
  summarize(
    mean = mean(richness),
    sd = sd(richness))

## Plot columns of exotic richness, 2016 seed bank vs. 2017, by zone
np_exotic_richness_may_col <- np_exotic_richness_may %>% 
  ggplot(aes(x = Plot, y = mean)) +
  geom_col(aes(fill = as.factor(Year)), position = "dodge") +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd, width = .5, fill = as.factor(Year)), position = position_dodge(width = 1)) +
  facet_wrap(~Replicate_Zone) +
labs(title = "Exotic Species Richness, by Zone", x = "Vernal Pool", y = "Species Richness", caption = "All pools decreased from 2016 seed bank to 2017; \n MS decreased in 2018, but RT & WT1 increased a lot in 2018") +
  theme_classic() +
  scale_fill_manual(values = c("darkgreen", "chartreuse3", "chartreuse2"), name = "Year", labels = c("2016 seed bank", "2017 monitoring", "2018 monitoring")) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_discrete(limits = c("RT", "PH1", "MS", "TP", "WT2", "CS", "WT1"))
np_exotic_richness_may_col

### Check diagnostic graphs
# plot(np_exotic_seed_bank_richnes_lm)
#diagnostic graphs look pretty good

### Calculate Pearson's r
# cor.test(np_exotic_seed_bank_richness$Year, np_exotic_seed_bank_richness$richness)
#r = -.59, medium correlation

## Kruskal-Wallis

### Check for normality
#np_exotic_seed_bank_richness_hist <- ggplot(np_exotic_seed_bank_richness) +
  geom_histogram(aes(x = richness))
#np_exotic_seed_bank_richnes_hist
####not normal

## Kruskal-Wallis test
#np_exotic_seed_bank_richness_kw <- kruskal.test(richness ~ Year, data = np_exotic_seed_bank_richness)
#np_exotic_seed_bank_richness_kw
###p = .001425, median exotic richness in 2016 seed bank different from 2017


## NMDS for 2016 seed bank vs. 2017
  
### Add Year column to seedbank_2016
seedbank_2016_year <- seedbank_2016 %>%
  add_column(Year = 2016)
  
### Convert data (only 2016 seed bank and May 2017 transition zone data) to NMDS format
may_2017_t <- np_percent_cover %>% 
  filter(Month == "MAY", Year == "2017", Replicate_Zone == "T") %>% 
  rename(Count = abundance) %>% 
  full_join(seedbank_2016_year) %>% 
  drop_na(Count) %>% 
  filter(Replicate_Zone == "T")
may_2017_t <- may_2017_t[-486,]
may_2017_t <- may_2017_t[-489,]  
may_2017_t <- may_2017_t %>%
  spread(Species_Full_Name, Count) %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  select(-Replicate, -Month, -Replicate_Number, -date, -month_number, -Native_Status, -Species, -Growth_Form, -Bloom_Start, -Bloom_End, -Dominant_Zone, -Max_Height, -Disturbance_Favored, -Wetland_Obligate, -Growth_Habit, -Family, -percent_cover)


### Perform NMDS for transition zone
may_2017_t_nmds <-  metaMDS(comm = select(may_2017_t, -Plot, -Replicate_Zone, -Year), distance = "bray", trace = FALSE, autotransform = FALSE)

### Make plot with polygons by Plot
ordiplot(may_2017_t_nmds, type = "point", label = TRUE)
ordihull(may_2017_t_nmds, groups = may_2017_t$Year, draw = "polygon", label = TRUE)

```


# 2. What environmental filters can be manipulated to decrease exotics?
## 2a. How does the seedbank compare to the aboveground cover?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 2a. How does the seedbank compare to the aboveground cover? (presence/absence data)

## Add Year column to seedbank_2016, add presence/absence column
seedbank_2016_pa <- seedbank_2016 %>%
  add_column(Year = 2016) %>% 
  group_by(Year, Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(
    max = max(Count)
  ) %>% 
  filter(max>0) %>% 
  add_column(present = 1)

## Add presence/absence column to 2017 percent cover data, join to 2016 seed bank
pa_2016_2017 <- np_percent_cover %>% 
  filter(Year == "2017") %>% 
  group_by(Year, Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(
    max = max(percent_cover)
  ) %>% 
  add_column(present = 1) %>% 
  full_join(seedbank_2016_pa) %>% 
  na.omit()

## Plot exotics in 2016 seed bank vs. 2017, by zone, for RT
rt_exotic_pa <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  filter(Plot == "RT") %>% 
  ggplot(aes(x = Species_Full_Name, y = present)) +
  geom_col(aes(fill = as.factor(Year))) +
  facet_wrap(~Replicate_Zone) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), label = c("2016 seed bank", "2017 monitoring"), name = "Year") +
    labs(title = "RT: Exotic Species in Seed Bank vs. Monitoring", x = "Species", y = "Presence", caption = "More exotics coming up in upland zone than are in seed bank") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
rt_exotic_pa

# Scatter comparing 2016 vs 2017 presence/absence
exotic_pa_scatter <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  spread(Year, present) %>% 
  rename(seventeen = "2017") %>% 
  rename(sixteen = "2016") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  ggplot(aes(x = sixteen, y = seventeen)) +
  geom_jitter(aes(color = Species_Full_Name, shape = Plot)) +
  facet_wrap(~Replicate_Zone) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), label = c("2016 seed bank", "2017 monitoring"), name = "Year") +
    labs(title = "Exotic Species in Seed Bank vs. Monitoring", x = "Species", y = "Presence", caption = "More exotics coming up in upland zone than are in seed bank") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
ggplotly(exotic_pa_scatter)


# Plot % cover for seed bank + monitoring species
exotic_pa_cover_both <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  spread(Year, present) %>% 
  rename(seventeen = "2017") %>% 
  rename(sixteen = "2016") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(both = seventeen + sixteen) %>% 
  filter(both == 2) %>% 
  ggplot(aes(y = max, x = Species_Full_Name)) +
  geom_point(aes(color = Plot)) +
  facet_wrap(~Replicate_Zone) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 5)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
exotic_pa_cover_both

# Plot % cover for just monitoring species
exotic_pa_cover <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  spread(Year, present) %>% 
  rename(seventeen = "2017") %>% 
  rename(sixteen = "2016") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(both = seventeen + sixteen) %>% 
  filter(seventeen == 1) %>% 
  filter(both == 1) %>% 
  ggplot(aes(y = max, x = Species_Full_Name)) +
  geom_point(aes(color = Plot)) +
  facet_wrap(~Replicate_Zone) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 45)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
exotic_pa_cover

```



## 2b. Does active management result in a different exotic community?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 2b. Does active management result in a different exotic community? (NMDS w/ old Del Sol data)

# Plot Del Sol percent cover
del_sol_cover <- del_sol %>% 
  full_join(metadata) %>% 
  drop_na(Plot) %>% 
  group_by(Plot, Species_Full_Name, Native_Status) %>% 
  summarize(
    cover = mean(cover)
  )

del_sol_cover_scatter <- ggplot(del_sol_cover, aes(x = Species_Full_Name, y = cover)) +
  geom_point(aes(color = Plot)) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 70)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~Native_Status)
del_sol_cover_scatter

# Plot Del Sol vs. NP 2017 percent cover, using mean of each species
del_sol_cover_col <-  np_percent_cover %>% 
  filter(Year == "2017") %>% 
  group_by(Year, Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(
    cover = max(percent_cover)
  ) %>% 
  full_join(del_sol_cover) %>% 
  drop_na(Native_Status) %>% 
  ggplot(aes(x = reorder(Species_Full_Name, -cover), y = cover)) +
  geom_col(aes(fill = as.factor(Year)), position = "dodge") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~Native_Status)
del_sol_cover_col
```







```{r echo = FALSE, message = FALSE, warning = FALSE}

# Trying to do non-independent stats tests for Cal-IPC poster -- friedman test

#time_friedman <- friedman.test(average_abundance ~ Plot | Creation_Year, data = np_annual_total_exotic_count)
#np_annual_total_exotic_count$Plot <- factor(np_annual_total_exotic_count$Plot)
#np_annual_total_exotic_count$Time_Since <- factor(np_annual_total_exotic_count$Time_Since)
#table(np_annual_total_exotic_count$Plot,np_annual_total_exotic_count$Year)

#wb <- aggregate(warpbreaks$breaks,
#                by = list(w = warpbreaks$wool,
#                          t = warpbreaks$tension),
#                FUN = mean)
#wb
#friedman.test(wb$x, wb$w, wb$t)
#friedman.test(x ~ w | t, data = wb)

```


```{r include = FALSE}

# Trying to do GLM for Cal-IPC poster

#time_glm <- glm(average_abundance~Time_Since, data = np_annual_total_exotic_count, family = c("quasipoisson"))
#summary(time_glm)

#time_glmer <- glmer(average_abundance~Time_Since + (1|Plot), data = np_annual_total_exotic_count, family = c("poisson"))
#AIC(time_glm, time_glmer)

#descdist(np_annual_total_exotic$average_percent_cover, discrete = FALSE)
#plot(fitdist(np_annual_total_exotic$average_percent_cover, distr = "beta"))

## Plot GLM using plot()
#time_x <- seq(1, 9, 1)
#time_y <- predict(time2_glm, list(Time_Since = time_x), type = "response")
#time_plot <- plot(np_annual_total_exotic_count$Time_Since, np_annual_total_exotic_count$average_abundance)
#lines(time_x, time_y)


###generalized linear model with binomial variance function and a logit link
#glm(count~site, weight=totalcount, data=speciesdata, family=binomial(link="logit"))


## Plot GLM using ggplot
#time_glm_plot <- np_annual_total_exotic_count %>% 
#  ggplot(aes(x = Time_Since, y = average_abundance)) +
#  geom_point(aes(color = Plot)) +
#  geom_smooth(method = "glm") +
#  facet_wrap(~Replicate_Zone)
#time_glm_plot

#GLS
#time_gls <- 

#Random effects LM
#lme(, random = ~ 1|Plot)

```


```{r echo = FALSE, message = FALSE, error = FALSE}

# Trying to do non-independent stats for Cal-IPC poster -- Beta regression

#glm(random = ~ 1|Plot)

```


```{r include = FALSE}

# Final graphs I was going to put into Cal-IPC 2019 poster: Compare total % cover of exotics (sum of max % cover of each species) over time since restoration

#Results: Significant increase in total exotic percent cover after year 5 -- due to increase in exotics in transition zone.


## Create new data frame with creation year of each pool
creation_year_df <- data.frame(Plot = c("PH1", "TP", "RT", "MS", "WT1", "WT2", "CS"), Creation_Year = c("2010", "2012", "2012", "2015", "2014", "2014", "2013"))

## Calculate total exotic % cover in 2017/2018, add time since restoration column
np_annual_total_exotic <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total = sum(percent_cover)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year)
  
## Boxplot of total % cover vs. time since
total_exotic_box <- np_annual_total_exotic %>% 
  drop_na() %>% 
  filter(Time_Since < 7) %>% 
  ggplot(aes(x = as.factor(Time_Since), y = total)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  geom_signif(comparisons = list(c("5", "6")), map_signif_level = TRUE) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Sig increase in median during year 6 (Dunn's Kruskal-Wallis Multiple Comparisons p = .0075)") +
  theme_classic() +
    scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100)) +
  theme(text = element_text(size=20))
total_exotic_box

## Boxplot of total % cover vs. time since, by zone
total_exotic_zone_box <- np_annual_total_exotic %>% 
  drop_na() %>% 
  filter(Time_Since < 7) %>% 
  ggplot(aes(x = as.factor(Time_Since), y = total)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  geom_signif(comparisons = list(c("5", "6")), map_signif_level = TRUE) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Sig increase in median during year 6 driven by sig incraese in transition zone (Dunn's Kruskal-Wallis Multiple Comparisons p = .0014)") +
  theme_classic() +
      scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 100)) +
  theme(text = element_text(size=20))
total_exotic_zone_box

## Exploratory graphs to determine ANOVA vs. KW
total_exotic_hist <- ggplot(np_annual_total_exotic, aes(x = total)) +
  geom_histogram()
total_exotic_hist
total_exotic_qq <- ggplot(np_annual_total_exotic, aes(sample = total)) +
  geom_qq()
total_exotic_qq

### Levene's test for equal variances
#np_annual_total_exotic$Time_Since <- as.character(np_annual_total_exotic$Time_Since)
#total_exotic_levene <- leveneTest(total ~ Time_Since, data = np_annual_total_exotic)
#total_exotic_levene
#p>.01, variances not equal

### Test of variances (is largest variance <4x smallest variance)
total_exotic_variances <- np_annual_total_exotic %>% 
  drop_na() %>% 
  group_by(Time_Since) %>% 
  summarize(variance = var(total))
#largest variance >4x smallest variance

## KW

#H0: Median total exotic percent cover across all years is equal
#HA: At least 2 medians differ significantly

np_annual_total_exotic$Time_Since <- as.factor(np_annual_total_exotic$Time_Since)
total_exotic_kw <- kruskal.test(total ~ Time_Since, data = np_annual_total_exotic)
total_exotic_kw

## Post-hoc Dunn's test
total_exotic_dunn <- dunnTest(total ~ Time_Since, data = np_annual_total_exotic)
print(total_exotic_dunn, dunn.test.results = TRUE) #5-6 is sig, p = .0075

##KW/Post-hoc per zone
np_annual_total_exotic_c <- np_annual_total_exotic %>% 
  filter(Replicate_Zone == "C")
total_exotic_c_kw <-kruskal.test(total ~ Time_Since, data = np_annual_total_exotic_c)
total_exotic_c_kw

np_annual_total_exotic_t <- np_annual_total_exotic %>% 
  filter(Replicate_Zone == "T")
total_exotic_t_kw <-kruskal.test(total ~ Time_Since, data = np_annual_total_exotic_t)
total_exotic_t_kw
total_exotic_dunn_t <- dunnTest(total ~ Time_Since, data = np_annual_total_exotic_t)
print(total_exotic_dunn_t, dunn.test.results = TRUE) #5-6 is sig (p = .0014)

np_annual_total_exotic_u <- np_annual_total_exotic %>% 
  filter(Replicate_Zone == "U")
total_exotic_u_kw <- kruskal.test(total ~ Time_Since, data = np_annual_total_exotic_u)
total_exotic_u_kw
total_exotic_dunn_u <- dunnTest(total ~ Time_Since, data = np_annual_total_exotic_u)
print(total_exotic_dunn_u, dunn.test.results = TRUE)

## ANOVA
#total_exotic_aov <- aov(total ~ Time_Since, data = np_annual_total_exotic)
#summary(total_exotic_aov)

## Post-hoc Tukey's HSD

#total_exotic_post_hoc <- TukeyHSD(total_exotic_aov)
#total_exotic_post_hoc #6-5 is sig


## AOV/Post-hoc per zone
#np_annual_total_exotic_c <- np_annual_total_exotic %>% 
#  filter(Replicate_Zone == "C")
#total_exotic_c_aov <- aov(total ~ Time_Since, data = np_annual_total_exotic_c)
#summary(total_exotic_c_aov)

#np_annual_total_exotic_t <- np_annual_total_exotic %>% 
#  filter(Replicate_Zone == "T")
#total_exotic_t_aov <- aov(total ~ Time_Since, data = np_annual_total_exotic_t)
#summary(total_exotic_t_aov)
#total_exotic_t_post_hoc <- TukeyHSD(total_exotic_t_aov)
#total_exotic_t_post_hoc #6-5 is sig

#np_annual_total_exotic_u <- np_annual_total_exotic %>% 
#  filter(Replicate_Zone == "U")
#total_exotic_u_aov <- aov(total ~ Time_Since, data = np_annual_total_exotic_u)
#summary(total_exotic_u_aov)
#total_exotic_u_post_hoc <- TukeyHSD(total_exotic_u_aov)
#total_exotic_u_post_hoc

```


