---
title: "NP Draft 6"
author: "Joanna Tang"
date: "4/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = FALSE}

# Loading packages, read in data
library(pairwiseAdonis)
library(kableExtra)
library(vegan)
library(car)
library(RColorBrewer)
library(reshape2)
library(kableExtra)
library(FSA)
library(plotly)
library(ggsignif)
library(lme4)
library(bestNormalize)
library(betareg)
library(nlme)
library(MASS)
library(fitdistrplus)
library(plotrix)
library(janitor)
library(ggpubr)
library(rstatix)
library(tidyverse)

np_percent_cover_master <- read_csv("np_percent_cover.csv")
metadata <- read_csv("metadata.csv")
np_2016_seedbank_master <- read_csv("np_2016_seedbank.csv")
np_2018_seed_bank_master <- read_csv("np.2018.seed.bank.csv")
biomass_master <- read_csv("combined_biomass.csv")
del_sol <- read_csv("delsol.csv")
noaa <- read_csv("1567450.csv")
np_pc_ord <- read_csv("pc-ord_nms_scores.csv")
np_community_matrix_variables <- read_csv("np_community_matrix_variables.csv")

```


```{r include = FALSE}

# Outline

# 0. Tidy up data

# 1. How does exotic community change over time?
## 1a. How does total exotic percent cover change with time since restoration (and interannual climate variability?)? (boxplots)
## 1b. Is there a threshold (# of years) after which exotic abundance increases? (linear model)
## 1c. Do pools converge on the same exotic community composition over time? (NMDS)

# 2. What environmental filters can be manipulated to decrease exotics?
## 2a. How does the seedbank compare to the aboveground cover? (presence/absence data)
## 2b. Does active management result in a different exotic community? (NMDS w/ old Del Sol data)
## 2c. Do certain exotics and natives occupy the same (phenological) niche? (NMDS, or scatterplot b/t 2 species)

# 3. What is the ideal monitoring plan?
## 3a. Same as 1b -- is there a threshold after which exotics increase (i.e., what's the ideal duration of monitoring?)?
## 3b. What is the ideal frequency/season of monitoring? (boxplots)


```


```{r include = FALSE}

# 0. Tidy up data

## Separate "Replicate" label into Replicate_Zone and Replicate_Number, join with species metadata, convert month and year to "date" column
np_percent_cover <- np_percent_cover_master %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3"))
np_percent_cover <- full_join(np_percent_cover, metadata) %>% 
  mutate(month_number = case_when(Month == "JANUARY" ~ 1, Month == "FEBRUARY" ~ 2, Month == "MARCH" ~ 3, Month == "APRIL" ~ 4, Month == "MAY" ~ 5, Month == "JUNE" ~ 6, Month == "JULY" ~ 7, Month == "AUGUST" ~ 8, Month == "SEPTEMBER" ~ 9, Month == "OCTOBER" ~ 10, Month == "NOVEMBER" ~ 11, Month == "DECEMBER" ~ 12)) %>% 
  mutate(date = paste(Year, month_number, "1", sep="-" )) %>% 
  dplyr::select(-X8)
np_percent_cover$date = as.Date(np_percent_cover$date)


## Convert np_2016_seedbank into tidy format, remove NAs, separate "Replicate" label into Replicate_Zone and Replicate_Number
seedbank_2016 <- gather(np_2016_seedbank_master, key = "Species", value = "Count", -c(Plot, Replicate)) %>% 
  na.omit(np_2016_seedbank_master) %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3")) %>% 
  filter(Count>0) %>% 
  full_join(metadata)

## Calculate total abundance of each species per Replicate for 2018 seed bank data, join to metadata
np_seedbank_2018 <- np_2018_seed_bank_master %>% 
  group_by(Plot, Replicate, Species) %>% 
  summarize(Count = sum(Count)) %>% 
  add_column(Year = 2018) %>% 
  full_join(metadata)


```

# 1. How does exotic community change over time?

## 1a. How does total exotic percent cover change with time since restoration?

Hypothesis: Exotics will be able to reinvade, causing increased exotic abundance and diversity with increased time since restoration.

Results: Significant increase in total exotic percent cover after year 6 -- due to increase in exotics in transition and upland zones.


```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1a. How does total exotic percent cover change with time since restoration (and interannual climate variability?)? (boxplots)


## Create new data frame with creation year of each pool
creation_year_df <- data.frame(Plot = c("PH1", "TP", "RT", "MS", "WT1", "WT2", "CS"), Creation_Year = c("2010", "2012", "2012", "2015", "2014", "2014", "2013"))

## Calculate total exotic % cover per quadrat, find max total exotic % cover per quadrat per year, add time since restoration column
np_annual_total_exotic <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  group_by(Year, Plot, Replicate, Replicate_Zone) %>% 
  summarize(max_annual_exotic_pc = max(total_exotic_pc)) %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year)

## Dataframe of median of the max annual total % cover of each pool
np_annual_total_exotic_medians <- np_annual_total_exotic %>% 
  group_by(Year, Plot, Time_Since) %>% 
  summarize(median = median(max_annual_exotic_pc))

## Boxplot of total exotic % cover
total_exotic_pc_box <- np_annual_total_exotic %>% 
  ggplot(aes(x = as.factor(Time_Since), y = max_annual_exotic_pc)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Increase in median after year 6 driven by sig incraese in transition and upland zones (Dunn's Kruskal-Wallis Multiple Comparisons p =)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
total_exotic_pc_box

## Boxplot of total exotic % cover, by zone
total_exotic_pc_zone_box <- np_annual_total_exotic %>% 
  ggplot(aes(x = as.factor(Time_Since), y = max_annual_exotic_pc)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  facet_wrap(~Replicate_Zone) +
  labs(title = "Total Percent Cover of Exotics, by zone", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Increase in median after year 6 driven by sig incraese in transition and upland zones (Dunn's Kruskal-Wallis Multiple Comparisons p =)") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
total_exotic_pc_zone_box

## Boxplot of total exotic % cover, by monitoring year -- everything's higher in 2019...
total_exotic_pc_year_box <- np_annual_total_exotic %>% 
  ggplot(aes(x = as.factor(Year), y = max_annual_exotic_pc)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  labs(title = "Total Percent Cover of Exotics", x = "Year", y = "Percent Cover", caption = "everything's higher in 2019...") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
total_exotic_pc_year_box




# Check for normality -- not normal
pc_hist <- ggplot(np_annual_total_exotic, aes(x = max_annual_exotic_pc)) +
  geom_histogram()
pc_qq <- ggplot(np_annual_total_exotic, aes(sample = max_annual_exotic_pc)) +
  geom_qq()

#### from original data analysis for Cal-IPC poster
## 95% CI around median
#lower CI = (n/2) - (1.96*sqrt(n))/2 th ranked value
#upper CI = 1 + (n/2) + (1.96*sqrt(n))/2 th ranked value

#n(3yrs) = 3
#lower CI = (3/2) - (1.96*sqrt(3/2)) = -.9 -> 1st ranked value = 2.38
#upper CI = 1 + (3/2) + (1.96*sqrt(3/2)) = 4.9005 -> 3rd ranked value -> 3.0

#n(4yrs) = 4
#lower CI = (4/2) - (1.96*sqrt(4/2)) = -.77 -> 1st ranked value = 2.73
#upper CI = 1 + (4/2) + (1.96*sqrt(4/2)) = 5.77 -> 4th ranked value -> 4.27

#n(5yrs) = 5
#lower CI = (5/2) - (1.96*sqrt(5/2)) = -.60 -> 1st ranked value = 2.65
#upper CI = 1 + (5/2) + (1.96*sqrt(5/2)) = 6.60 -> 5th ranked value -> 11.533

#n(6yrs) = 3
#lower CI = (3/2) - (1.96*sqrt(3/2)) = -.9 -> 1st ranked value = 4.90
#upper CI = 1 + (3/2) + (1.96*sqrt(3/2)) = 4.9005 -> 3rd ranked value -> 17.86

#n(7yrs) = 3
#lower CI = (3/2) - (1.96*sqrt(3/2)) = -.9 -> 1st ranked value = 8.60
#upper CI = 1 + (3/2) + (1.96*sqrt(3/2)) = 4.9005 -> 3rd ranked value -> 19.12

exotic_pc_ci_df <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  filter(Time_Since >2) %>% 
  filter(Time_Since <8) %>% 
  group_by(Time_Since) %>% 
  summarize(
    median = median(total_exotic_pc),
    max = max(total_exotic_pc),
    min = min(total_exotic_pc)
  )

exotic_pc_ci_zone_df <- np_percent_cover %>%
  filter(Native_Status == "E") %>% 
  filter(Year > 2016) %>% 
  group_by(date, Year, Month, Plot, Replicate, Replicate_Zone) %>% 
  summarize(total_exotic_pc = sum(percent_cover)) %>% 
  drop_na() %>% 
  full_join(creation_year_df) %>% 
  mutate(Creation_Year = as.numeric(as.character(Creation_Year))) %>% 
  mutate(Time_Since = Year - Creation_Year) %>% 
  filter(Time_Since >2) %>% 
  filter(Time_Since <8) %>% 
  group_by(Time_Since, Replicate_Zone) %>% 
  summarize(
    median = median(total_exotic_pc),
    max = max(total_exotic_pc),
    min = min(total_exotic_pc)
  )

#### from original data analysis for Cal-IPC poster
#n(3yrs) = 23
#lower CI = (23/2) - (1.96*sqrt(23/2)) = 4.85 -> 4th ranked value = 1.33
#upper CI = 1 + (23/2) + (1.96*sqrt(23/2)) = 19.15 -> 20th ranked value -> 3.50

#n(4yrs) = 30
#lower CI = (30/2) - (1.96*sqrt(30/2)) = 7.41 -> 7th ranked value = 2.00
#upper CI = 1 + (30/2) + (1.96*sqrt(30/2)) = 23.59 ->24th ranked value -> 4.33

#n(5yrs) = 34
#lower CI = (34/2) - (1.96*sqrt(34/2)) = 9.92 -> 9th ranked value = 2.50
#upper CI = 1 + (34/2) + (1.96*sqrt(34/2)) = 26.08 -> 26th ranked value -> 8.67

#n(6yrs) = 23
#lower CI = (23/2) - (1.96*sqrt(23/2)) = 4.85 -> 4th ranked value = 2.00
#upper CI = 1 + (23/2) + (1.96*sqrt(23/2)) = 19.15 -> 20th ranked value -> 15.00

#n(7yrs) = 22
#lower CI = (22/2) - (1.96*sqrt(22/2)) = 4.50 -> 4th ranked value = 2.50
#upper CI = 1 + (22/2) + (1.96*sqrt(22/2)) = 18.50 -> 18th ranked value -> 25.33

## Column graph of total % cover vs. time since, w/ 95% CI
total_exotic_col <- exotic_pc_ci_df %>% 
  ggplot(aes(x = as.factor(Time_Since), y = median)) +
  geom_col() +
  geom_errorbar(aes(ymin = min, ymax = max, width = .2)) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Sig increase in median during year 6 (Dunn's Kruskal-Wallis Multiple Comparisons p = )") +
  theme_classic() +
    scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 25)) +
  theme(text = element_text(size=20))
#total_exotic_col

## Column graph of total % cover vs. time since, w/ 95% CI, by zone
total_avg_exotic_zone_col <- exotic_pc_ci_zone_df %>% 
   ggplot(aes(x = as.factor(Time_Since), y = median)) +
  geom_col() +
  geom_errorbar(aes(ymin = min, ymax = max, width = .2)) +
  facet_wrap(~Replicate_Zone) +
  geom_signif(comparisons = list(c("5", "6")), map_signif_level = TRUE) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Percent Cover", caption = "Sig increase in median during year 6 driven by sig incraese in transition zone (Dunn's Kruskal-Wallis Multiple Comparisons p = )") +
  theme_classic() +
      scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 30)) +
  theme(text = element_text(size=20))
#total_avg_exotic_zone_col

```


## 1b. Is there a threshold (# of years) after which exotic abundance increases?

Hypothesis: We will see increased exotic species percent cover and diversity with increased time since restoration due to reinvasion.

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 1b. Is there a threshold (# of years) after which exotic abundance increases? (linear model)

# Scatterplot of median of annual max total % exotic cover
np_annual_total_exotic_medians_scatter <- np_annual_total_exotic_medians %>% 
  ggplot(aes(x = Time_Since, y = median, group = Plot)) +
  geom_point(aes(color = Plot)) +
  geom_smooth(method = lm, aes(x = Time_Since, group = NULL), se = TRUE, size = .5) +
  labs(title = "Total Percent Cover of Exotics", x = "Time Since Restoration (years)", y = "Median Percent Cover (%)", caption = "High variation; WT2, PH1, & esp WT1 and RT increased") +
  theme_classic() +
  scale_color_brewer(palette = "Dark2", name = "Vernal Pool") +
  scale_y_continuous(expand = c(0,0))
np_annual_total_exotic_medians_scatter

```



## 1c. Do pools converge on the same exotic community composition over time?

```{r include = FALSE}

# Exploratory graphs using vegan

## Community matrix
np_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:141)



## How speciose are my communities?
np_richness <- specnumber(np_community_matrix)

### Does species richness differ b/t pools?
np_richness_aov <- aov(np_richness ~ Plot, data = np_community_matrix_variables)
#summary(np_richness_aov)
#p = .779, F = .078
#no sig difference of richness b/t pools

#### Richness, by plot
np_richness_box <- np_richness %>% 
  enframe() %>% 
  full_join(np_community_matrix_variables, by = c("name" = "replicate_date")) %>% 
  ggplot(aes(x = as.factor(Plot), y = value)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  labs(title = "Species Richness, by plot", x = "Pool", y = "Richness", caption = "No difference in species richness across plots (ANOVA p = 0.779, F = 0.078))") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
#np_richness_box




## How diverse are my communities?
np_shannon <- diversity(np_community_matrix)

### Does diversity differ b/t pools?
np_shannon_aov <- aov(np_shannon ~ Plot, data = np_community_matrix_variables)
#summary(np_shannon_aov)
#p = .082, F = .775
#no sig difference of diversity b/t pools

#### Shannon's Diversity, by plot
np_shannon_box <- np_shannon %>% 
  enframe() %>% 
  full_join(np_community_matrix_variables, by = c("name" = "replicate_date")) %>% 
  ggplot(aes(x = as.factor(Plot), y = value)) +
  geom_boxplot() +
  geom_jitter(width = .2, alpha = .4) +
  labs(title = "Species Diversity, by plot", x = "Pool", y = "Shannon's Index", caption = "No difference in Shannon's Diversity across plots (ANOVA p = 0.082, F = 0.775))") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0))
#np_shannon_box






## How different are my communities in species composition? (perMANOVA)
np_permanova <- adonis(np_community_matrix ~ Plot, data = np_community_matrix_variables)
#p = .051, F = 2.1151

### NMDS
np_nmds <- metaMDS(np_community_matrix)
#stressplot(np_nmds)
#Generally, an average stress of > 0.2 is an indication that the ordination didn’t do a good job of representing community structure. However, NMDS stress increases as sample size increases, so if you have many samples (as this dataset does) your ordination will likely exhibit a lot of stress

np_nmds_scores <- scores(np_nmds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>% 
  full_join(np_community_matrix_variables)

np_nmds_fit <- envfit(np_nmds, np_community_matrix, perm = 999) #envfit() takes the output of metaMDS() and the species matrix you created

np_nmds_fit_pvals <- np_nmds_fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") #extract p-values for each species

np_nmds_fit_coords <- np_nmds_fit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., np_nmds_fit_pvals, by = "species") %>% 
  filter(pvals == 0.001) #extract coordinates for species, only keep species with p-val = 0.001

### NMDS, by plot
np_nmds_plot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2, color = as.factor(Plot), shape = as.factor(Plot))) +
  geom_point() +
  stat_ellipse(linetype = 2, size = 1) +
  theme_classic() +
  labs(title = "NMDS, by plot", caption = "No difference in community across plots (perANOVA p = 0.051, F = 2.11))")
#np_nmds_plot






## How is community structure related to specific environmental variables? (CCA)

np_cca <- cca(np_community_matrix ~ creation_year + do_mg + depth_in + salinity + area_m2 + month, data = np_community_matrix_variables)
np_cca_baseplot <- plot(np_cca)
#get the scaling right for the vectors (blue in the base plot). To do this, you can store the base plot in an object, run that object, and extract the attr(,"arrow.mul") value from the summary that pops up in the console.
#attr(,"arrow.mul") = 27.11853

np_cca_vectors <- as.matrix(scores(np_cca, display = "bp", scaling = "species")*27.11853
) %>% 
  as.data.frame() #extract vectors

np_cca_site_coords <- scores(np_cca, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>% 
  full_join(np_community_matrix_variables) #extract site coordinates

np_cca_species_coords <- scores(np_cca, display = "species") %>% 
  as.data.frame() #extract species coordinates

#### CCA, with environmental biplot
np_cca_biplot <- ggplot(np_cca_site_coords) +
  geom_point(aes(x = CCA1, y = CCA2, color = as.factor(Plot)), shape = 19, size = 2, alpha = 0.8) +
  coord_fixed() +
  geom_segment(data = np_cca_vectors, aes(x = 0, y = 0, xend = CCA1, yend = CCA2), arrow = arrow(length = unit(0.2, "cm"))) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  geom_point(data = np_cca_species_coords, aes(x = CCA1, y = CCA2), shape = 17, size = 2, color = "slateblue") +
  scale_x_continuous(limits = c(-12, 10)) +
  scale_y_continuous(limits = c(-3, 12)) +
  geom_text(data = np_cca_vectors, aes(x = CCA1, y = CCA2, label = rownames(np_cca_vectors)), nudge_x = 0.3, nudge_y = 0.3) +
  labs(title = "Canonical Correspondence Analysis")
np_cca_biplot

```


```{r include = FALSE}

# NMDS with pc-ord NMS1 data
creation_year_numbered_df <- creation_year_df %>% mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19"))

## Create new data frame with creation year of each pool, depth of each pool, water quality
pool_factors_df <- data.frame(Plot = c("1", "7", "4", "9", "14", "16", "19"), depth_in = c("16", "18", "9", "16.5", "16", "17", "15"), area_m2 = c("173.13", "346.79", "85.28", "160.07", "432.89", "424.18", "359.38"), do_mg = c("3.73", "4.825", "4.635", "8.22", "2.925", "3.25", "4.44"), do_sat = c("38.4", "49.75", "49.15", "85.2", "28.6", "33.5", "44.25"), conductivity = c("1156", "91.25", "236", "674", "214", "288.95", "491.8"), salinity = c(".7", ".05", ".15", ".4", ".15", "1.1", ".3")) %>% 
  full_join(creation_year_numbered_df) %>% 
  mutate(Creation_Year = as.character(Creation_Year)) %>% 
  mutate(Creation_Year = as.numeric(Creation_Year))

## Community matrix dataframe with all variables
np_nmds_df <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  full_join(pool_factors_df) %>% 
  mutate(age = Year - Creation_Year) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0)))

## Used the above community matrix and variables to run NMS in PC-ORD with Scott Cooper 11/13/2020

### Dataframe of NMS1 scores
nms1_df <- np_pc_ord %>% 
  mutate(plot = case_when(plot == "1" ~ "PH1", plot == "7" ~ "TP", plot == "4" ~ "RT", plot == "9" ~ "MS", plot == "14" ~ "WT1", plot == "16" ~ "WT2", plot == "19" ~ "CS")) %>% 
  mutate(replicate_zone = case_when(replicate_zone == "1" ~ "C", replicate_zone == "2" ~ "T", replicate_zone == "3" ~ "U")) %>% 
  mutate(replicate = paste(plot, replicate_zone, sep = "_")) %>% 
  dplyr::select(replicate, plot, replicate_zone, month_year, nms1) %>% 
  pivot_wider(names_from = month_year, values_from = nms1) %>% 
  pivot_longer(4:31, names_to = "month_year", values_to = "nms1")

### Exploratory graphs, checking assumptions
nms1_hist <- nms1_df %>% 
  ggplot() +
  geom_histogram(aes(x = nms1)) +
  facet_wrap(~month_year)
#kind of normal-ish -- skewed left-ish
nms1_qq <- nms1_df %>% 
  ggplot() +
  geom_qq(aes(sample = nms1)) +
  facet_wrap(~month_year)
#kind of normal-ish -- skewed left-ish
nms1_normality <- nms1_df %>% 
  group_by(month_year) %>% 
  shapiro_test(nms1)

### Repeated measures ANOVA for NMS1
nms1_rmanova <- anova_test(data = nms1_df, nms1 ~ month_year, wid = replicate)
#get_anova_table(nms1_rmanova)
#p << .001, F(488) = 6.717, ges = .256
#### Pairwise t-tests
nms1_t_tests <- pairwise_t_test(data = nms1_df, nms1 ~ month_year, paired = TRUE, p.adjust.method = "bonferroni")



### NMS1 by plot
#### ANOVA
nms1_anova_plot <- anova_test(data = nms1_df, nms1 ~ plot, wid = replicate)
#get_anova_table(nms1_anova_plot)
#p << .001, F = 5.848, ges = .06
#### Pairwise t-tests
nms1_t_plot <- pairwise_t_test(data = nms1_df, nms1 ~ plot, paired = TRUE, p.adjust.method = "bonferroni")
nms1_t_plot_df <- as.data.frame(pairwise.t.test(nms1_df$nms1, nms1_df$plot)$p.value)
#MS-PH1 p < .001, MS-RT p = .00594, MS-TP p = .210, MS-WT1 p = .062, MS-WT2 p < .001, PH1-CS p = .0898

#### NMDS, with each replicate and species names
### Community matrix dataframe
nms1_plot_community_matrix <- np_nmds_df %>% 
  clean_names() %>% 
    mutate(plot = case_when(plot == "1" ~ "PH1", plot == "7" ~ "TP", plot == "4" ~ "RT", plot == "9" ~ "MS", plot == "14" ~ "WT1", plot == "16" ~ "WT2", plot == "19" ~ "CS")) %>% 
  mutate(replicate_zone = case_when(replicate_zone == "1" ~ "C", replicate_zone == "2" ~ "T", replicate_zone == "3" ~ "U")) %>% 
  mutate(replicate = paste(plot, replicate_zone, sep = "_"))

### Community matrix dataframe
np_community_matrix <- np_nmds_df %>% 
  clean_names() %>% 
    mutate(plot = case_when(plot == "1" ~ "PH1", plot == "7" ~ "TP", plot == "4" ~ "RT", plot == "9" ~ "MS", plot == "14" ~ "WT1", plot == "16" ~ "WT2", plot == "19" ~ "CS")) %>% 
  mutate(replicate_zone = case_when(replicate_zone == "1" ~ "C", replicate_zone == "2" ~ "T", replicate_zone == "3" ~ "U")) %>% 
  mutate(replicate = paste(plot, replicate_zone, sep = "_"))
np_community_matrix <- np_nmds_df[, -c(1:5)]
np_community_matrix <- np_community_matrix[, -c(2:10)]
np_community_matrix <- np_community_matrix %>% 
  column_to_rownames(var = "replicate_date")

### NMDS, with each replicate and species names
np_nmds <- metaMDS(np_community_matrix, distance = "bray", k=2, wascores = TRUE)
np_nmds_scores <- np_nmds$species
plot(np_nmds)
orditorp(np_nmds, "species")
orditorp(np_nmds,display="sites",cex=.25,air=0.01)


```



```{r echo = FALSE, message = FALSE, warning = FALSE}
# 1c. Do pools converge on the same exotic community composition over time? (NMDS)

## Exotic community matrix
exotic_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "E") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:70)

## Exotic community matrix variables
exotic_community_matrix_variables <- exotic_community_matrix %>% 
  rownames_to_column(var = "replicate_date") %>% 
  select(replicate_date) %>% 
  inner_join(np_community_matrix_variables)

#exotic_nmds <- metaMDS(exotic_community_matrix)
#stressplot(exotic_nmds)
#Generally, an average stress of > 0.2 is an indication that the ordination didn’t do a good job of representing community structure. However, NMDS stress increases as sample size increases, so if you have many samples (as this dataset does) your ordination will likely exhibit a lot of stress
#stress = .00015

exotic_nmds_scores <- scores(np_nmds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>% 
  full_join(exotic_community_matrix_variables) %>% 
  drop_na(Plot)

exotic_nmds_fit <- envfit(exotic_nmds, exotic_community_matrix, perm = 999) #envfit() takes the output of metaMDS() and the species matrix you created

exotic_nmds_fit_pvals <- exotic_nmds_fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") #extract p-values for each species

exotic_nmds_fit_coords <- exotic_nmds_fit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., np_nmds_fit_pvals, by = "species") %>% 
  filter(pvals == 0.001) #extract coordinates for species, only keep species with p-val = 0.001


### NMDS for exotics, by plot, with species biplot
exotic_nmds_plot_species_biplot <- ggplot(exotic_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = Plot)) +
  stat_ellipse(aes(color = Plot)) +
  geom_segment(data = exotic_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = exotic_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS of exotic community, by plot", caption = "PH1 has different community (post-hoc pairwise perMANOVA p = 0.021) \n WT1 also has slightly different community (post-hoc pairwise perMANOVA p = 0.021) \ MS also has slightly different community (post-hoc pairwise perMANOVA p - .021)")
exotic_nmds_plot_species_biplot

### perMANOVA, by plot
exotic_plot_permanova <- adonis(exotic_community_matrix ~ Plot, data = exotic_community_matrix_variables)
#p = .001
###Post-hoc test
exotic_plot_pairwise <- pairwise.adonis(exotic_community_matrix, exotic_community_matrix_variables$Plot)
#PH1-WT1 p = .021, PH1-WT2 p = .063, PH1-CS p = .021, PH1-RT p = .021, PH1-TP p = .084, PH1-MS p = .021, WT1-CS p = .042, WT1-RT p = .021, WT1-MS p = .021, WT1-CS p = .021, WT1-RT p = .021, WT2-MS p = .021, CS-RT p = .021, CS-MS p = .021, RT-MS p = .021, MS-TP p = .021



### NMDS for exotics, by zone, with species biplot
exotic_nmds_zone_species_biplot <- ggplot(exotic_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = Replicate_Zone)) +
  stat_ellipse(aes(color = Replicate_Zone)) +
  geom_segment(data = exotic_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = exotic_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS of exotic community, by zone", caption = "Sig difference between center and transition zones (post-hoc pairwise perMANOVA p = 0.003) \n Sig difference between center and upland zones (post-hoc pairwise perMANOVA p = 0.003) \n Sig difference between transition and upland zones (post-hoc pairwise perMANOVA p = 0.003)")
exotic_nmds_zone_species_biplot

### perMANOVA, by zone
exotic_zone_permanova <- adonis(exotic_community_matrix ~ Replicate_Zone, data = exotic_community_matrix_variables)
#p = .001
###Post-hoc test
exotic_zone_pairwise <- pairwise.adonis(exotic_community_matrix, exotic_community_matrix_variables$Replicate_Zone)
#C-T p = .003, C-U p = .003, T-U p = .003




### NMDS for exotics, by age, with species biplot
exotic_nmds_age_species_biplot <- ggplot(exotic_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = as.factor(age))) +
  stat_ellipse(aes(color = as.factor(age))) +
  geom_segment(data = exotic_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = exotic_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS of exotic community, by pool age", caption = "Sig difference in exotic community across age (perMANOVA p = 0.001)) -- community changes at year 4")
exotic_nmds_age_species_biplot

### perMANOVA, by age
exotic_age_permanova <- adonis(exotic_community_matrix ~ age, data = exotic_community_matrix_variables)
#p = .001
###Post-hoc test
exotic_age_pairwise <- pairwise.adonis(exotic_community_matrix, exotic_community_matrix_variables$age)
#6-2 p = .028, 6-3 p = .056, 2-3 p = .028, 2-4 p = .028, 2-5 p = .028, 2-7 p = .028, 2-8 p = .028, 2-9 p = .028, 3-5 p = .028, 3-8 p = .028, 3-9 p = .028, 4-7 p = .028, 5-8 p = .028



### NMDS for exotics, by month, with species biplot
exotic_nmds_month_species_biplot <- ggplot(exotic_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = month)) +
  stat_ellipse(aes(color = month)) +
  geom_segment(data = exotic_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = exotic_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS of exotic community, by monitoring month", caption = "Slight sig differences in exotic community between months")
#exotic_nmds_month_species_biplot

### perMANOVA, by month
exotic_month_permanova <- adonis(exotic_community_matrix ~ month, data = exotic_community_matrix_variables)
#p = .001
###Post-hoc test
exotic_month_pairwise <- pairwise.adonis(exotic_community_matrix, exotic_community_matrix_variables$month)
#p = .055 for some Dec, Jan, Feb, March, April, May, 



### NMDS for exotics, by year, with species biplot
exotic_nmds_year_species_biplot <- ggplot(exotic_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = as.factor(Year))) +
  stat_ellipse(aes(color = as.factor(Year))) +
  geom_segment(data = exotic_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = exotic_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS of exotic community, by montoring year", caption = "Sig difference in exotic community between years (perMANOVA p = 0.001)")
exotic_nmds_year_species_biplot

### perMANOVA, by month
exotc_year_permanova <- adonis(exotic_community_matrix ~ Year, data = exotic_community_matrix_variables)
#p = .001
###Post-hoc test
exotic_year_pairwise <- pairwise.adonis(exotic_community_matrix, exotic_community_matrix_variables$Year)
#all p = .006

```
# 2. Is the community changing?

## 2a. Is the whole community changing?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 2a. Is the whole community changing?

## Community matrix
np_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:141)

#np_nmds <- metaMDS(np_community_matrix)
#stressplot(np_nmds)
#Generally, an average stress of > 0.2 is an indication that the ordination didn’t do a good job of representing community structure. However, NMDS stress increases as sample size increases, so if you have many samples (as this dataset does) your ordination will likely exhibit a lot of stress

np_nmds_scores <- scores(np_nmds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("replicate_date") %>% 
  full_join(np_community_matrix_variables)

np_nmds_fit <- envfit(np_nmds, np_community_matrix, perm = 999) #envfit() takes the output of metaMDS() and the species matrix you created

np_nmds_fit_pvals <- np_nmds_fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") #extract p-values for each species

np_nmds_fit_coords <- np_nmds_fit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., np_nmds_fit_pvals, by = "species") %>% 
  filter(pvals == 0.001) #extract coordinates for species, only keep species with p-val = 0.001



### NMDS, by plot, with species biplot
np_nmds_plot_species_biplot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = Plot)) +
  stat_ellipse(aes(color = Plot)) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species)) +
  theme_classic() +
  labs(title = "NMDS, by plot", caption = "PH1 has different community (post-hoc pairwise perMANOVA p = 0.021) \n RT also has slightly different community (post-hoc pairwise perMANOVA p = 0.021)")
np_nmds_plot_species_biplot

### perMANOVA, by plot
np_plot_permanova <- adonis(np_community_matrix ~ Plot, data = np_community_matrix_variables)
#p = .001
###Post-hoc test
np_plot_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$Plot)
#PH1-RT p = .021, PH1-TP p = .021, PH1-MS p = .021, PH1-WT1 p = .021, PH1-WT2 p = .021, PH1-CS p = .021, RT-MS p = .021, RT-WT2 p = .021, RT-CS p = .021, TP-WT2 p = .063



### NMDS, by zone, with species biplot
np_nmds_zone_species_biplot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = Replicate_Zone)) +
  stat_ellipse(aes(color = Replicate_Zone)) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species)) +
  theme_classic() +
  labs(title = "NMDS, by zone", caption = "Sig difference between center and transition zones (post-hoc pairwise perMANOVA p = 0.059) \n Sig difference between center and upland zones (post-hoc pairwise perMANOVA p = 0.044)")
np_nmds_zone_species_biplot

### perMANOVA, by zone
np_zone_permanova <- adonis(np_community_matrix ~ Replicate_Zone, data = np_community_matrix_variables)
#p = .068
###Post-hoc test
np_zone_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$Replicate_Zone)
#C-T p = .059, C-U p = .044

### NMDS, by age, with species biplot
np_nmds_age_species_biplot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = as.factor(age))) +
  stat_ellipse(aes(color = as.factor(age))) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS, by pool age", caption = "Sig difference in community across age (perMANOVA p = 0.001)) -- community changes at year 6")
np_nmds_age_species_biplot

### perMANOVA, by age
np_age_permanova <- adonis(np_community_matrix ~ age, data = np_community_matrix_variables)
#p = .001
###Post-hoc test
np_age_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$age)
#7-8 p = .086, 7-9 p = .048, 7-6 p = .001, 7-5 p = .001, 7-4 p = .001, 7-3 p = .001, 7-2 p = .001, 8-6 p = .001, 8-5 p = .001, 8-4 p = .001, 8-3 p = .001, 8-2 p = .001, 9-6 p = .001, 9-5 p = .001, 9-4 p = .001, 9-3 p = .001, 9-2 p = .001, 6-5 p = .073, 6-4 p = .002, 6-3 p = .015, 6-2 p = .011



### NMDS, by month, with species biplot
np_nmds_month_species_biplot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = month)) +
  stat_ellipse(aes(color = month)) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS, by monitoring month", caption = "No sig difference in community across monitoring months (perMANOVA p = 0.866))")
np_nmds_month_species_biplot

### perMANOVA, by month
np_month_permanova <- adonis(np_community_matrix ~ month, data = np_community_matrix_variables)
#p = .866




### NMDS, by year, with species biplot
np_nmds_year_species_biplot <- ggplot(np_nmds_scores, aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  geom_point(aes(color = as.factor(Year))) +
  stat_ellipse(aes(color = as.factor(Year))) +
  geom_segment(data = np_nmds_fit_coords, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = np_nmds_fit_coords, aes(label = species), alpha = .75) +
  theme_classic() +
  labs(title = "NMDS, by montoring year", caption = "Slight difference in community \n between 2017-2019 (post-hoc pairwise perMANOVA p = 0.066) \n and between 2016-2018 (post-hoc pairwise perMANOVA p = 0.084)")
np_nmds_year_species_biplot

### perMANOVA, by month
np_year_permanova <- adonis(np_community_matrix ~ Year, data = np_community_matrix_variables)
#p = .099
###Post-hoc test
np_year_pairwise <- pairwise.adonis(np_community_matrix, np_community_matrix_variables$Year)
#2017-2019 p = .066, 2016-2018 p = .084

```

## 2b. Is the native community changing?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 2a. Is the native community changing?

## Native community matrix
native_community_matrix <- np_percent_cover %>% 
  filter(Replicate_Zone != "NA") %>% 
  filter(Native_Status == "N") %>% 
  mutate(Replicate_Zone = case_when(Replicate_Zone == "C" ~ "1", Replicate_Zone == "T" ~ "2", Replicate_Zone == "U" ~ "3")) %>% 
  mutate(Plot = case_when(Plot == "PH1" ~ "1", Plot == "TP" ~ "7", Plot == "RT" ~ "4", Plot == "MS" ~ "9", Plot == "WT1" ~ "14", Plot == "WT2" ~ "16", Plot == "CS" ~"19")) %>% 
  group_by(date, month_number, Year, Plot, Replicate_Zone, Species) %>% 
  summarize(mean_pc = mean(percent_cover)) %>% 
  mutate(replicate_date = paste(Plot, Replicate_Zone, date, sep = "-")) %>% 
  mutate(month_year = paste(Year, month_number, sep = ".")) %>% 
  pivot_wider(names_from = Species, values_from = mean_pc) %>% 
  mutate_all(funs(replace_na(.,0))) %>% 
  column_to_rownames("replicate_date") %>% 
  select(7:68)

## Native community matrix variables
native_community_matrix_variables <- native_community_matrix %>% 
  rownames_to_column(var = "replicate_date") %>% 
  select(replicate_date) %>% 
  inner_join(np_community_matrix_variables)

#native_nmds <- metaMDS(native_community_matrix, maxit = 4000)
#stressplot(native_nmds)
#Generally, an average stress of > 0.2 is an indication that the ordination didn’t do a good job of representing community structure. However, NMDS stress increases as sample size increases, so if you have many samples (as this dataset does) your ordination will likely exhibit a lot of stress

```

# 2. What environmental filters can be manipulated to decrease exotics?
## 2a. How does the seedbank compare to the aboveground cover?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 2a. How does the seedbank compare to the aboveground cover? (presence/absence data)

## Add Year column to seedbank_2016, add presence/absence column
seedbank_2016_pa <- seedbank_2016 %>%
  add_column(Year = 2016) %>% 
  group_by(Year, Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(
    max = max(Count)
  ) %>% 
  filter(max>0) %>% 
  add_column(present = 1)

## Add presence/absence column to 2017 percent cover data, join to 2016 seed bank
pa_2016_2017 <- np_percent_cover %>% 
  filter(Year == "2017") %>% 
  group_by(Year, Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(
    max = max(percent_cover)
  ) %>% 
  add_column(present = 1) %>% 
  full_join(seedbank_2016_pa) %>% 
  na.omit()

## Plot exotics in 2016 seed bank vs. 2017, by zone, for RT
rt_exotic_pa <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  filter(Plot == "RT") %>% 
  ggplot(aes(x = Species_Full_Name, y = present)) +
  geom_col(aes(fill = as.factor(Year))) +
  facet_wrap(~Replicate_Zone) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), label = c("2016 seed bank", "2017 monitoring"), name = "Year") +
    labs(title = "RT: Exotic Species in Seed Bank vs. Monitoring", x = "Species", y = "Presence", caption = "More exotics coming up in upland zone than are in seed bank") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
rt_exotic_pa

# Scatter comparing 2016 vs 2017 presence/absence
exotic_pa_scatter <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  spread(Year, present) %>% 
  rename(seventeen = "2017") %>% 
  rename(sixteen = "2016") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  ggplot(aes(x = sixteen, y = seventeen)) +
  geom_jitter(aes(color = Species_Full_Name, shape = Plot)) +
  facet_wrap(~Replicate_Zone) +
  scale_fill_manual(values = c("darkgreen", "chartreuse4"), label = c("2016 seed bank", "2017 monitoring"), name = "Year") +
    labs(title = "Exotic Species in Seed Bank vs. Monitoring", x = "Species", y = "Presence", caption = "More exotics coming up in upland zone than are in seed bank") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
ggplotly(exotic_pa_scatter)


# Plot % cover for seed bank + monitoring species
exotic_pa_cover_both <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  spread(Year, present) %>% 
  rename(seventeen = "2017") %>% 
  rename(sixteen = "2016") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(both = seventeen + sixteen) %>% 
  filter(both == 2) %>% 
  ggplot(aes(y = max, x = Species_Full_Name)) +
  geom_point(aes(color = Plot)) +
  facet_wrap(~Replicate_Zone) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 5)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
exotic_pa_cover_both

# Plot % cover for just monitoring species
exotic_pa_cover <- pa_2016_2017 %>% 
  filter(Native_Status == "E") %>% 
  spread(Year, present) %>% 
  rename(seventeen = "2017") %>% 
  rename(sixteen = "2016") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(both = seventeen + sixteen) %>% 
  filter(seventeen == 1) %>% 
  filter(both == 1) %>% 
  ggplot(aes(y = max, x = Species_Full_Name)) +
  geom_point(aes(color = Plot)) +
  facet_wrap(~Replicate_Zone) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 45)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
exotic_pa_cover

```



## 2b. Does active management result in a different exotic community?

```{r echo = FALSE, message = FALSE, warning = FALSE}

# 2b. Does active management result in a different exotic community? (NMDS w/ old Del Sol data)

# Plot Del Sol percent cover
del_sol_cover <- del_sol %>% 
  full_join(metadata) %>% 
  drop_na(Plot) %>% 
  group_by(Plot, Species_Full_Name, Native_Status) %>% 
  summarize(
    cover = mean(cover)
  )

del_sol_cover_scatter <- ggplot(del_sol_cover, aes(x = Species_Full_Name, y = cover)) +
  geom_point(aes(color = Plot)) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 70)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~Native_Status)
del_sol_cover_scatter

# Plot Del Sol vs. NP 2017 percent cover, using mean of each species
del_sol_cover_col <-  np_percent_cover %>% 
  filter(Year == "2017") %>% 
  group_by(Year, Plot, Replicate_Zone, Species_Full_Name, Native_Status) %>% 
  summarize(
    cover = max(percent_cover)
  ) %>% 
  full_join(del_sol_cover) %>% 
  drop_na(Native_Status) %>% 
  ggplot(aes(x = reorder(Species_Full_Name, -cover), y = cover)) +
  geom_col(aes(fill = as.factor(Year)), position = "dodge") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~Native_Status)
del_sol_cover_col

```


