---
title: "np_nmds"
author: "Joanna Tang"
date: "12/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

# Loading packages, read in data
library(tidyverse)
library(knitr)
library(kableExtra)
library(vegan)
np_2016_seedbank_master <- read_csv("np_2016_seedbank.csv")
np_percent_cover <- read_csv("np_percent_cover.csv")
metadata <- read_csv("metadata.csv")

```

```{r include = FALSE}

# Separate "Replicate" label into Replicate_Zone and Replicate_Number, join with species metadata, convert month and year to "date" column
percent_cover <- np_percent_cover %>% 
  mutate(Replicate_Zone = case_when(Replicate == "C1" ~ "C", Replicate == "C2" ~ "C", Replicate == "C3" ~ "C", Replicate == "T1" ~ "T", Replicate == "T2" ~ "T", Replicate == "T3" ~ "T", Replicate == "U1" ~ "U", Replicate == "U2" ~ "U", Replicate == "U3" ~ "U")) %>% 
  mutate(Replicate_Number = case_when(Replicate == "C1" ~ "1", Replicate == "C2" ~ "2", Replicate == "C3" ~ "3", Replicate == "T1" ~ "1", Replicate == "T2" ~ "2", Replicate == "T3" ~ "3", Replicate == "U1" ~ "1", Replicate == "U2" ~ "2", Replicate == "U3" ~ "3"))
percent_cover <- full_join(percent_cover, metadata) %>% 
  mutate(month_number = case_when(Month == "JANUARY" ~ 1, Month == "FEBRUARY" ~ 2, Month == "MARCH" ~ 3, Month == "APRIL" ~ 4, Month == "MAY" ~ 5, Month == "JUNE" ~ 6, Month == "JULY" ~ 7, Month == "AUGUST" ~ 8, Month == "SEPTEMBER" ~ 9, Month == "OCTOBER" ~ 10, Month == "NOVEMBER" ~ 11, Month == "DECEMBER" ~ 12)) %>% 
  mutate(date = paste(Year, month_number, "1", sep="-" ))
percent_cover$date = as.Date(percent_cover$date)

```


# NMDS for Percent Cover April 2017

```{r}

# NMDS for Percent Cover April 2017

# Convert data (only April 2017 central zone data) to NMDS format
april_2017 <- percent_cover %>% 
  filter(Month == "APRIL", Year == "2017", Replicate_Zone == "C") %>% 
  spread(Species, percent_cover) %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  select(-Replicate, -Month, -Year, -abundance, -Replicate_Number, -date, -month_number, -Native_Status, -Species_Full_Name)

# Perform NMDS
april_2017_nmds <-  metaMDS(comm = select(april_2017, -Plot, -Replicate_Zone), distance = "bray", trace = FALSE, autotransform = FALSE)


# Make plot of NMDS, colored by Plot
MDS_xy <- data.frame(april_2017_nmds$points)
MDS_xy$Plot <- april_2017$Plot
## This makes a dataframe of MDS points organized by Plot

april_2017_nmds_graph <- ggplot(MDS_xy, aes(MDS1, MDS2, color = Plot)) +
  geom_point(alpha = .3)
april_2017_nmds_graph

# Make plot with polygons by Plot
ordiplot(april_2017_nmds, type = "point")
ordihull(april_2017_nmds, groups = april_2017$Plot, draw = "polygon", label = TRUE)

```

Plots all have overlapping polygons -- they're very similar (no matter the month). Zones also have overlapping polygons.  The most difference is seen between zones in April 2017.

# NMDS for Seed Bank 2016

```{r}

# NMDS for Seed Bank 2016

# Remove NAs from data, use only central zone data
seedbank_2016 <- na.omit(np_2016_seedbank_master) %>% 
  filter(Replicate == "C1" | Replicate == "C2" | Replicate == "C3")

# Perform NMDS
seedbank_2016_nmds <- metaMDS(comm = select(seedbank_2016, -Plot, -Replicate), distance = "bray", trace = FALSE, autotransform = FALSE)

# Make plot of NMDS, colored by Plot
MDS_xy <- data.frame(seedbank_2016_nmds$points)
MDS_xy$Plot <- seedbank_2016$Plot
## This makes a dataframe of MDS points organized by Plot

seedbank_2016_nmds_graph <- ggplot(MDS_xy, aes(MDS1, MDS2, color = Plot)) +
  geom_point(alpha = .3)
seedbank_2016_nmds_graph

# Make plot of polygons by Plot
ordiplot(seedbank_2016_nmds, type = "n")
ordihull(seedbank_2016_nmds, groups = seedbank_2016$Plot, draw = "polygon", label = TRUE)

```

Seed bank shows overlapping polygons for Plots -- they're all very similar overall.  However, only plotting the central zone shows dissimilarity.  Only plotting the transition zone also shows dissimilarity, while only plotting the upland zone shows similarity (upland invaded zone is similar).